---
title: "AAS Nuclei GM Volume Analysis - LC & nbM"
output: html_document
---

```{r setup}
library(lmPerm)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(broom)

setwd('~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native')

# Define ROIs of interest
roi_labels <- c(485, 486, 491, 492)
roi_names <- c("485" = "L_LC", "486" = "R_LC", "491" = "R_nbM", "492" = "L_nbM")
```

## Load in AAS nuclei volume data (native space, 1mm)

```{r load_aas_volumes}
subject_ids_file <- '~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/subject_ids.txt'
subject_ids <- readLines(subject_ids_file)
data_list <- list()

# Loop through each subject ID and load the corresponding .csv file
for (sub_id in subject_ids) {
  csv_file <- paste0(
    "~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/",
    sub_id, "_GM_aas_nuclei_volumes.csv"
  )
  
  if (file.exists(csv_file)) {
    data <- read.csv(csv_file,header=TRUE)
    data$SubjectID <- sub_id
    data_list[[sub_id]] <- data
  } else {
    message(paste("File not found for subject:", sub_id))
  }
}

# Combine all dataframes into a single dataframe
combined_data <- bind_rows(data_list)
head(combined_data)

# Filter to only the 4 ROIs of interest
combined_data <- combined_data %>%
  filter(Label %in% roi_labels) %>%
  mutate(ROI_Name = roi_names[as.character(Label)])
```

```{r pivot_volumes}
# Probabilistic GM Volume (wide format)
ProbGM_Volume <- combined_data %>%
  select(SubjectID, ROI_Name, ProbGM_Volume_mm3) %>%
  pivot_wider(names_from = ROI_Name, values_from = ProbGM_Volume_mm3)

# Binary Mask Volume (wide format)
Binary_Volume <- combined_data %>%
  select(SubjectID, ROI_Name, BinaryMask_Volume_mm3) %>%
  pivot_wider(names_from = ROI_Name, values_from = BinaryMask_Volume_mm3)

# Mean GM Probability - all voxels (wide format)
MeanGM_All <- combined_data %>%
  select(SubjectID, ROI_Name, MeanGM_AllVoxels) %>%
  pivot_wider(names_from = ROI_Name, values_from = MeanGM_AllVoxels)

# Mean GM Probability - non-zero voxels (wide format)
MeanGM_NonZero <- combined_data %>%
  select(SubjectID, ROI_Name, MeanGM_NonZero) %>%
  pivot_wider(names_from = ROI_Name, values_from = MeanGM_NonZero)
```

## Load demographic and clinical data

```{r load_demographics}
cog_data <- read.csv("~/onedrive_usyd/Postdoc_Rob/Analysis/Demographic_Data/mri_scanner_cog.csv")
subject_data <- read.csv("~/onedrive_usyd/Postdoc_Rob/Analysis/Demographic_Data/subject_data_IL810_mri.csv")

# Standardize subject_data$sub_id to match "sub-00X" format
subject_data$sub_id <- sprintf("sub-%03d", subject_data$sub_id)
subject_data <- subject_data[-88, ] #need to remove sub-164 as it is not in the pc data
subject_data <- subject_data[-c(30,31), ] #remove sub-049 and sub-050 as they are not in the pc data
# Find unmatched subjects
unmatched_subjects <- setdiff(subject_ids, subject_data$sub_id)
print(paste("Unmatched subjects:", paste(unmatched_subjects, collapse = ", ")))

# Filter volume data to remove unmatched subjects
ProbGM_Volume <- ProbGM_Volume %>% filter(!(SubjectID %in% unmatched_subjects))
Binary_Volume <- Binary_Volume %>% filter(!(SubjectID %in% unmatched_subjects))
MeanGM_All <- MeanGM_All %>% filter(!(SubjectID %in% unmatched_subjects))
MeanGM_NonZero <- MeanGM_NonZero %>% filter(!(SubjectID %in% unmatched_subjects))
```

## Load TIV data

```{r load_tiv}
data_list_tiv <- list()

for (sub_id in subject_ids) {
  csv_file <- paste0(
    "~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/Volume_TIV/",
    sub_id, "_tiv_volumes_brainmsk.csv"
  )
  
  if (file.exists(csv_file)) {
    data2 <- read.csv(csv_file)
    data2$SubjectID <- sub_id
    data_list_tiv[[sub_id]] <- data2
  } else {
    message(paste("TIV file not found for subject:", sub_id))
  }
}

Volume_TIV <- bind_rows(data_list_tiv) %>%
  select(SubjectID, Volume_mm3) %>%
  filter(!(SubjectID %in% unmatched_subjects))
```

## Descriptive Statistics - Check GM probability values in brainstem

```{r descriptive_check}
# This is critical: check whether GM segmentation is reliable in these nuclei
cat("=== Mean GM Probability (All Voxels) - Summary per ROI ===\n")
combined_data %>%
  filter(!(SubjectID %in% unmatched_subjects)) %>%
  group_by(ROI_Name) %>%
  summarise(
    mean_prob = mean(MeanGM_AllVoxels, na.rm = TRUE),
    sd_prob = sd(MeanGM_AllVoxels, na.rm = TRUE),
    min_prob = min(MeanGM_AllVoxels, na.rm = TRUE),
    max_prob = max(MeanGM_AllVoxels, na.rm = TRUE),
    n = n()
  ) %>%
  print()

cat("\n=== Probabilistic GM Volume (mm3) - Summary per ROI ===\n")
combined_data %>%
  filter(!(SubjectID %in% unmatched_subjects)) %>%
  group_by(ROI_Name) %>%
  summarise(
    mean_vol = mean(ProbGM_Volume_mm3, na.rm = TRUE),
    sd_vol = sd(ProbGM_Volume_mm3, na.rm = TRUE),
    min_vol = min(ProbGM_Volume_mm3, na.rm = TRUE),
    max_vol = max(ProbGM_Volume_mm3, na.rm = TRUE),
    n = n()
  ) %>%
  print()
```

## Statistical Analysis - Probabilistic GM Volume by Delirium Group

```{r ttest_probGM}
# Merge with group information
ProbGM_with_groups <- ProbGM_Volume %>%
  left_join(subject_data %>% select(sub_id, bin_delirium), by = c("SubjectID" = "sub_id"))

# T-tests for each ROI
results_probGM <- ProbGM_with_groups %>%
  pivot_longer(cols = -c(SubjectID, bin_delirium), names_to = "ROI", values_to = "Volume") %>%
  group_by(ROI) %>%
  summarise(
    mean_group0 = mean(Volume[bin_delirium == 0], na.rm = TRUE),
    mean_group1 = mean(Volume[bin_delirium == 1], na.rm = TRUE),
    t_statistic = t.test(Volume ~ bin_delirium)$statistic,
    p_value = t.test(Volume ~ bin_delirium)$p.value
  ) %>%
  mutate(significance = case_when(
    p_value < 0.001 ~ "***",
    p_value < 0.01 ~ "**",
    p_value < 0.05 ~ "*",
    TRUE ~ "ns"
  ))

cat("=== T-test Results: Probabilistic GM Volume ===\n")
print(results_probGM)

library(kableExtra)

results_probGM %>%
  mutate(
    mean_group0 = round(mean_group0, 2),
    mean_group1 = round(mean_group1, 2),
    t_statistic = round(t_statistic, 3),
    p_value = formatC(p_value, format = "e", digits = 3)
  ) %>%
  kable(
    col.names = c("ROI", "Mean (No Delirium)", "Mean (Delirium)", 
                   "t-statistic", "p-value", "Sig."),
    caption = "T-test Results: Probabilistic GM Volume",
    align = "lccccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  row_spec(which(results_probGM$p_value < 0.05), bold = TRUE, color = "red")

```

```{r ttest_binary}
# Also test binary mask volume (atlas region volume)
Binary_with_groups <- Binary_Volume %>%
  left_join(subject_data %>% select(sub_id, bin_delirium), by = c("SubjectID" = "sub_id"))

results_binary <- Binary_with_groups %>%
  pivot_longer(cols = -c(SubjectID, bin_delirium), names_to = "ROI", values_to = "Volume") %>%
  group_by(ROI) %>%
  summarise(
    mean_group0 = mean(Volume[bin_delirium == 0], na.rm = TRUE),
    mean_group1 = mean(Volume[bin_delirium == 1], na.rm = TRUE),
    t_statistic = t.test(Volume ~ bin_delirium)$statistic,
    p_value = t.test(Volume ~ bin_delirium)$p.value
  ) %>%
  mutate(significance = case_when(
    p_value < 0.001 ~ "***",
    p_value < 0.01 ~ "**",
    p_value < 0.05 ~ "*",
    TRUE ~ "ns"
  ))

cat("=== T-test Results: Binary Mask Volume ===\n")
print(results_binary)
results_binary %>%
  mutate(
    mean_group0 = round(mean_group0, 2),
    mean_group1 = round(mean_group1, 2),
    t_statistic = round(t_statistic, 3),
    p_value = formatC(p_value, format = "e", digits = 3)
  ) %>%
  kable(
    col.names = c("ROI", "Mean (No Delirium)", "Mean (Delirium)", 
                   "t-statistic", "p-value", "Sig."),
    caption = " T-test Results: Binary Mask Volume",
    align = "lccccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  row_spec(which(results_probGM$p_value < 0.05), bold = TRUE, color = "red")

```

## Statistical Analysis - Ratio of Probabilistic GM Volume / TIV

```{r ratio_analysis}
# Calculate ratio of ProbGM Volume / TIV
ProbGM_with_TIV <- ProbGM_Volume %>%
  left_join(Volume_TIV, by = "SubjectID")

Ratio_output <- ProbGM_with_TIV %>%
  mutate(across(
    c(L_LC, R_LC, R_nbM, L_nbM),
    ~ . / Volume_mm3,
    .names = "Ratio_{.col}"
  )) %>%
  select(SubjectID, starts_with("Ratio_"))

# Merge with group information
Ratio_with_groups <- Ratio_output %>%
  left_join(subject_data %>% select(sub_id, bin_delirium), by = c("SubjectID" = "sub_id"))

# T-tests on ratios
results_ratio <- Ratio_with_groups %>%
  pivot_longer(cols = -c(SubjectID, bin_delirium), names_to = "ROI", values_to = "Ratio") %>%
  group_by(ROI) %>%
  summarise(
    mean_group0 = mean(Ratio[bin_delirium == 0], na.rm = TRUE),
    mean_group1 = mean(Ratio[bin_delirium == 1], na.rm = TRUE),
    t_statistic = t.test(Ratio ~ bin_delirium)$statistic,
    p_value = t.test(Ratio ~ bin_delirium)$p.value
  ) %>%
  mutate(significance = case_when(
    p_value < 0.001 ~ "***",
    p_value < 0.01 ~ "**",
    p_value < 0.05 ~ "*",
    TRUE ~ "ns"
  ))

cat("=== T-test Results: ProbGM Volume / TIV Ratio ===\n")
print(results_ratio)

results_ratio %>%
  mutate(
    mean_group0 = formatC(mean_group0, format = "e", digits = 3),
    mean_group1 = formatC(mean_group1, format = "e", digits = 3),
    t_statistic = round(t_statistic, 3),
    p_value = formatC(p_value, format = "e", digits = 3),
    ROI = gsub("Ratio_", "", ROI)
  ) %>%
  kable(
    col.names = c("ROI", "Mean (No Delirium)", "Mean (Delirium)",
                   "t-statistic", "p-value", "Sig."),
    caption = "T-test Results: Probabilistic GM Volume / TIV Ratio",
    align = "lccccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(which(results_ratio$p_value < 0.05), bold = TRUE, color = "red")

```

## Logistic Regression - AAS nuclei volumes and TIV predicting delirium

```{r logistic_regression}
# Calculate bilateral averages
# Calculate bilateral averages
logistic_data <- ProbGM_Volume %>%
  mutate(
    LC_avg = (L_LC + R_LC) / 2,
    nbM_avg = (L_nbM + R_nbM) / 2
  ) %>%
  left_join(Volume_TIV, by = "SubjectID") %>%
  left_join(subject_data %>% select(sub_id, bin_delirium), by = c("SubjectID" = "sub_id"))

# Loop through each region
regions <- c("L_LC", "R_LC", "L_nbM", "R_nbM", "LC_avg", "nbM_avg")
logistic_models <- list()
logistic_summaries <- list()

for (region in regions) {
  formula <- as.formula(paste("bin_delirium ~", region, "+ Volume_mm3"))
  model <- glm(formula, data = logistic_data, family = binomial)
  logistic_models[[region]] <- model
  logistic_summaries[[region]] <- summary(model)
}

# Tidy results
results_logres <- map_dfr(logistic_models, ~ broom::tidy(.), .id = "Region")

results_logres %>%
  mutate(
    estimate = round(estimate, 4),
    std.error = round(std.error, 4),
    statistic = round(statistic, 3),
    p.value = formatC(p.value, format = "e", digits = 3)
  ) %>%
  kable(
    col.names = c("Region", "Term", "Estimate", "Std. Error",
                   "z-statistic", "p-value"),
    caption = "Logistic Regression Results: ROI Volume + TIV ~ Delirium",
    align = "llcccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

capture.output(logistic_summaries, 
               file = "~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/logistic_GM_volume_native_aas_nuclei_TIV_bin_del.txt")
```

## Plotting - Box plots by delirium group

```{r boxplot_probGM, fig.width=10, fig.height=6}
roi_order <- c("L_LC", "R_LC", "L_nbM", "R_nbM")

ProbGM_long <- ProbGM_with_groups %>%
  pivot_longer(cols = -c(SubjectID, bin_delirium), names_to = "ROI", values_to = "Volume") %>%
  left_join(results_probGM %>% select(ROI, significance), by = "ROI") %>%
  mutate(ROI = factor(ROI, levels = roi_order))

p_box <- ggplot(ProbGM_long, aes(x = as.factor(bin_delirium), y = Volume, fill = as.factor(bin_delirium))) +
  geom_boxplot(outlier.shape = 21, outlier.size = 2) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
  facet_wrap(~ ROI, scales = "free_y", nrow = 1) +
  geom_text(
    data = ProbGM_long %>%
      group_by(ROI, significance) %>%
      summarise(y_pos = max(Volume, na.rm = TRUE) * 1.05, .groups = "drop") %>%
      distinct(),
    aes(x = 1.5, y = y_pos, label = significance),
    inherit.aes = FALSE, size = 5, hjust = 0.5
  ) +
  scale_fill_manual(values = c("0" = "#f9a340", "1" = "#7fb6cc"),
                    labels = c("0" = "No Delirium", "1" = "Delirium")) +
  labs(
    title = "Probabilistic GM Volume by Delirium Group - AAS Nuclei",
    x = "Delirium Group",
    y = expression("Probabilistic GM Volume (mm"^3*")"),
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "bottom"
  )

print(p_box)
ggsave(
  filename = "~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/ProbGM_Volume_AAS_Nuclei_Delirium_Group.svg",
  plot = p_box, width = 10, height = 6, dpi = 300
)
```

```{r boxplot_ratio, fig.width=10, fig.height=6}
ratio_order <- c("Ratio_L_LC", "Ratio_R_LC", "Ratio_L_nbM", "Ratio_R_nbM")

Ratio_long <- Ratio_with_groups %>%
  pivot_longer(cols = -c(SubjectID, bin_delirium), names_to = "ROI", values_to = "Ratio") %>%
  left_join(results_ratio %>% select(ROI, significance), by = "ROI") %>%
  mutate(ROI = factor(ROI, levels = ratio_order))

p_ratio <- ggplot(Ratio_long, aes(x = as.factor(bin_delirium), y = Ratio, fill = as.factor(bin_delirium))) +
  geom_boxplot(outlier.shape = 21, outlier.size = 2) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
  facet_wrap(~ ROI, scales = "free_y", nrow = 1) +
  geom_text(
    data = Ratio_long %>%
      group_by(ROI, significance) %>%
      summarise(y_pos = max(Ratio, na.rm = TRUE) * 1.05, .groups = "drop") %>%
      distinct(),
    aes(x = 1.5, y = y_pos, label = significance),
    inherit.aes = FALSE, size = 5, hjust = 0.5
  ) +
  scale_fill_manual(values = c("0" = "#f9a340", "1" = "#7fb6cc"),
                    labels = c("0" = "No Delirium", "1" = "Delirium")) +
  labs(
    title = "GM Volume / TIV Ratio by Delirium Group - AAS Nuclei",
    x = "Delirium Group",
    y = "ProbGM Volume / TIV Ratio",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "bottom"
  )

print(p_ratio)

ggsave(
  filename = "~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/ProbGM_Volume_TIV_Ratio_AAS_Nuclei_Delirium_Group.svg",
  plot = p_ratio, width = 10, height = 6, dpi = 300
)
```

## Scatter plots - Volume vs TIV by group

```{r scatter_tiv, fig.width=10, fig.height=6}
scatter_order <- c("L_LC", "R_LC", "L_nbM", "R_nbM", "LC_avg", "nbM_avg")

scatter_data <- logistic_data %>%
  pivot_longer(cols = c(L_LC, R_LC, L_nbM, R_nbM, LC_avg, nbM_avg),
               names_to = "ROI", values_to = "AvgVolume") %>%
  mutate(ROI = factor(ROI, levels = scatter_order))

p_scatter <- ggplot(scatter_data, aes(x = Volume_mm3, y = AvgVolume, color = as.factor(bin_delirium))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, linetype = "dashed") +
  facet_wrap(~ ROI, scales = "free_y") +
  scale_color_manual(values = c("0" = "#f9a340", "1" = "#7fb6cc"),
                     labels = c("0" = "No Delirium", "1" = "Delirium")) +
  labs(
    title = "AAS Nuclei GM Volume vs TIV by Delirium Group",
    x = "Total Intracranial Volume (TIV, mm³)",
    y = expression("Probabilistic GM Volume (mm"^3*")"),
    color = "Group"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "bottom"
  )

print(p_scatter)
ggsave(filename = "~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/ProbGM_Volume_AAS_vs_TIV_Scatter.svg", plot = p_scatter, width = 10, height = 6, dpi = 300)
```

## Mean GM Probability check - are brainstem segmentations reliable?

```{r gm_prob_check, fig.width=10, fig.height=5}
roi_labels <- c(485, 486, 491, 492)
roi_names <- c("L_LC", "R_LC", "R_nbM", "L_nbM")
roi_order <- c("L_LC", "R_LC", "L_nbM", "R_nbM")

MeanGM_long <- combined_data %>%
  filter(Label %in% roi_labels) %>%
  filter(!(SubjectID %in% unmatched_subjects)) %>%
  mutate(ROI_Name = factor(
    roi_names[match(Label, roi_labels)],
    levels = roi_order
  )) %>%
  left_join(subject_data %>% select(sub_id, bin_delirium), by = c("SubjectID" = "sub_id"))

p_prob <- ggplot(MeanGM_long, aes(x = ROI_Name, y = MeanGM_AllVoxels, fill = as.factor(bin_delirium))) +
  geom_boxplot(outlier.shape = 21, outlier.size = 2) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5, position = position_jitterdodge()) +
  geom_hline(yintercept = 0.3, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 0.5, y = 0.32, label = "Reliability threshold (0.3)",
           hjust = 0, color = "red", size = 3.5) +
  scale_fill_manual(values = c("0" ="#f9a340", "1" = "#7fb6cc"),
                    labels = c("0" = "No Delirium", "1" = "Delirium")) +
  labs(
    title = "Mean GM Probability in AAS Nuclei Masks",
    subtitle = "Values consistently below 0.3 suggest cortical segmentation is unreliable in brainstem",
    x = "ROI",
    y = "Mean GM Probability (all mask voxels)",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 12)
  )

print(p_prob)
ggsave(filename ="~/onedrive_usyd/Postdoc_Rob/Analysis/LC_volume/1mm_native/MeanGM_Probability_AAS_Nuclei_Check.svg", plot = p_prob, width = 10, height = 5, dpi = 300)
```
