---
title: "Blood biomarker analysis pre/post-op"
output: github_document
---

```
```{r load libraries}
#library(lmPerm)
library(broom)
library(ggplot2)
```
## Function Code for Logistic Regression
```{r}
## Define the function
run_glm_model <- function(subject_data, data_fc, response_var, covariates) {
  # Ensure variable names are valid
  names(subject_data) <- make.names(names(subject_data))
  names(data_fc) <- make.names(names(data_fc))
  response_var <- make.names(response_var)
  covariates <- make.names(covariates)
  
  # Extract the response variable
  response <- subject_data[[response_var]]
  
  # Combine the data
  data_comb <- cbind(response, data_fc, subject_data[covariates])
  data_model_df <- data.frame(data_comb)
  # Define predictor variables
  predictor_vars <- names(data_comb)[names(data_comb) != "response"]

  # Define the function to run the logistic regression model with permutation
  run_glm <- function(var, data_comb) {
    formula <- as.formula(paste("response ~", var, "*", paste(covariates, collapse = " + ")))
    model <- glm(formula, data = data_comb, family = binomial)
    return(model)
  }
  
  results <- lapply(predictor_vars, function(var) {
    glm <- run_glm(var, data_comb)
    summary_model <- summary(glm)
    marginal_effects <- predict(glm, data_comb,type = "response")
    return(list(model = glm, summary = summary_model, marginal_effects = marginal_effects))
  })
  names(results) <- predictor_vars  # Assign names to the results list
  # Extract beta coefficients and p-values
  coef_pval_list <- lapply(results, function(res) {
    coef_table <- summary(res$model)$coefficients
    data.frame(
      Variable = rownames(coef_table),
      Beta = coef_table[, "Estimate"],
      PValue = coef_table[, "Pr(>|z|)"],
      OddsRatio = exp(coef_table[, "Estimate"]) # Calculate odds ratio
    )
  })
  #Extract AIC Values
    aic_list <- lapply(results, function(res) {
      aic <- AIC(res$model)
      return(aic)
    })
  names(aic_list) <- predictor_vars  # Assign names to the AIC list
  # Combine all results into a single data frame
  coef_pval_df <- do.call(rbind, coef_pval_list)
  
  return(list(results = results, coef_pval_df = coef_pval_df,data_model_df = data_model_df, aic_list = aic_list))
}
```
# function logistic regression for gaussian
```{r}
run_linear_reg <- function(subject_data, data_fc, response_var, covariates) {
  # Ensure variable names are valid
  names(subject_data) <- make.names(names(subject_data))
  names(data_fc) <- make.names(names(data_fc))
  response_var <- make.names(response_var)
  covariates <- make.names(covariates)
  
  # Extract the response variable
  response <- subject_data[[response_var]]
  
  # Combine the data
  data_comb <- cbind(response, data_fc, subject_data[covariates])
  data_model_df <- data.frame(data_comb)
  predictor_vars <- names(data_comb)[names(data_comb) != "response"]
  
  # Define the function to run the linear regression model
  run_glm <- function(var, data_comb) {
    formula <- as.formula(paste("response ~", var, "*", paste(covariates, collapse = " + ")))
    model <- glm(formula, data = data_comb, family = gaussian)
    return(model)
  }
  
  results <- lapply(predictor_vars, function(var) {
    glm <- run_glm(var, data_comb)
    summary_model <- summary(glm)
    fitted_vals <- predict(glm, data_comb, type = "response")
    return(list(model = glm, summary = summary_model, fitted = fitted_vals))
  })
  names(results) <- predictor_vars
  
  # Extract coefficients, standard errors, t-values, and p-values
  coef_pval_list <- lapply(results, function(res) {
    coef_table <- summary(res$model)$coefficients
    data.frame(
      Variable = rownames(coef_table),
      Estimate = coef_table[, "Estimate"],
      StdError = coef_table[, "Std. Error"],
      tValue = coef_table[, "t value"],
      PValue = coef_table[, "Pr(>|t|)"]
    )
  })
  
  # Extract AIC values
  aic_list <- lapply(results, function(res) {
    aic <- AIC(res$model)
    return(aic)
  })
  names(aic_list) <- predictor_vars
  
  coef_pval_df <- do.call(rbind, coef_pval_list)
  
  return(list(results = results, coef_pval_df = coef_pval_df, data_model_df = data_model_df, aic_list = aic_list))
}
```

## Plot on the Brain Significant Logs Odds Ratio
```{r}
# Plot the significant coefficients on the brain
plot_brain_data <- function(data, filename) {
  library(ggplot2)
  library(ggseg)
  library(ggsegSchaefer)
  library(dplyr)
  library(stringr)
  
  # Ensure the data has a 'label' column
  if (!"label" %in% colnames(data)) {
    stop("The data must contain a 'label' column.")
  }
  
  # Prepare the data for plotting
  data_plot <- data %>%
    mutate(label = ifelse(str_detect(label, "_LH"), paste0("lh_", label), paste0("rh_", label)))
  
  # Print the data frame to verify
  View(data_plot)
  
  # Create the plot
  p <- ggplot(data_plot, aes(fill = SignificantCoefficient)) +
    geom_brain(atlas = schaefer17_400, show.legend = TRUE) +
    scale_fill_gradient2(low = "#C03B2B", mid = "white", high = "#0C7D79", midpoint = 0) +
    theme_void() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.key.size = unit(1.5, "cm"))
  
  # Print the plot
  print(p)
  #setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Manuscripts/AAS_Manuscript/Figures/Supplementary_Figs')
 setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/lm_model/") 
  # Save the plot as an SVG file
  ggsave(paste0(filename, ".svg"), plot = p, width = 10, height = 10)
}

plot_brain_data_oddsratio <- function(data, filename) {
  library(ggplot2)
  library(ggseg)
  library(ggsegSchaefer)
  library(dplyr)
  library(stringr)
  
  # Ensure the data has a 'label' column
  if (!"label" %in% colnames(data)) {
    stop("The data must contain a 'label' column.")
  }
  
  # Prepare the data for plotting
  data_plot <- data %>%
    mutate(label = ifelse(str_detect(label, "_LH"), paste0("lh_", label), paste0("rh_", label)))
  
  # Print the data frame to verify
  View(data_plot)
  
  # Create the plot
  p <- ggplot(data_plot, aes(fill = SignificantCoefficient)) +
    geom_brain(atlas = schaefer17_400, show.legend = TRUE) +
    scale_fill_gradient2(low = "#CA1C6A", mid = "white", high = "#173F6C", midpoint = 1, 
               limits = c(0, max(data$SignificantCoefficient, na.rm = TRUE))) +
    theme_void() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.key.size = unit(1.5, "cm"))
  
  # Print the plot
  print(p)
  #setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Manuscripts/AAS_Manuscript/Figures/Supplementary_Figs')
 #setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/lm_model/") 
  # Save the plot as an SVG file
  ggsave(paste0(filename, ".svg"), plot = p, width = 10, height = 10)
}
```
```{r}
plot_avg_fc_vs_var <- function(
  fc_data,
  y_vector,
  group_vector,
  y_label = "Y Variable",
  plot_title = "Relationship between Avg PC and MMN Avg Amplitude",
  filename = "avg_fc_vs_yvar.svg"
) {
  # Calculate average FC
  avg_fc <- rowMeans(fc_data, na.rm = TRUE)
  
  # Correlation
  cor_avg_fc <- cor.test(y_vector, avg_fc, use = "complete.obs")
  
  # Prepare data frame
  plot_df <- data.frame(
    avg_fc = avg_fc,
    y_var = y_vector,
    group = factor(group_vector, levels = c(0, 1), labels = c("Non-Delirious", "Delirious"))
  )
  colours <- c("Non-Delirious" = "#f9a340", "Delirious" = "#7fb6cc")
  
  # Plot
  library(ggplot2)
  p <- ggplot(plot_df, aes(x = avg_fc, y = y_var, color = group)) +
    geom_point(size = 3) +
    geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "black") +
    annotate("text", x = max(avg_fc, na.rm = TRUE) * 0.8, y = max(plot_df$y_var, na.rm = TRUE) * 0.9, 
             label = paste("r =", round(cor_avg_fc$estimate, 2), 
                           "\np =", signif(cor_avg_fc$p.value, 3)), 
             color = "black", size = 5) +
    labs(
      title = plot_title,
      x = "Avg FC",
      y = y_label,
      color = "Delirium Group"
    ) +
    scale_color_manual(values = colours) +
    theme_minimal()
  
  print(p)
  ggsave(filename = filename, plot = p, width = 8, height = 6, dpi = 300)
}

# Example usage:
# plot_avg_fc_vs_var(fc_data, subject_data_mmn_update$peak_time, subject_data_mmn_update$bin_delirium, y_label = "Peak Time", plot_title = "Relationship between Avg NBM FC and Peak Time", filename = "avg_nbm_fc_vs_mmn_peaktime.svg")
```

```{r load data}
# Load the data
setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data')
data <- read.csv("NPQ_2024-10-04_Lennertz_CNS120_Results_REDCap_IDs.csv",header=TRUE, stringsAsFactors = FALSE)

#separate the pre and postop data
baseline_data <- data[data$timepoint == "Baseline", ]
POD1_data <- data[data$timepoint == "POD1", ]
```

## plot all the data
```{r plot all data}
# scatter plot 
# Prepare data frame
match_sub<- match(baseline_data$sample_id, POD1_data$sample_id)
baseline_filter <- !is.na(match_sub) & match_sub > 0
baseline_data <- baseline_data[baseline_filter, ]

POD1_filter <- !is.na(match_sub) & match_sub > 0
POD1_data <- POD1_data[POD1_filter, ]

combined_data <- rbind(
    data.frame(
    timepoint = rep("Baseline", length(baseline_data$sample_id)),
    patient_id = baseline_data$sample_id,
    pTau181 = baseline_data$pTau.181,
    pTau217 = baseline_data$pTau.217,
    pTau231 = baseline_data$pTau.231,
    NEFL = baseline_data$NEFL,
    MAPT = baseline_data$MAPT
    ),
    data.frame(
    timepoint = rep("POD 1", length(POD1_data$sample_id)),
    patient_id = POD1_data$sample_id,
    pTau181 = POD1_data$pTau.181,
    pTau217 = POD1_data$pTau.217,
    pTau231 = POD1_data$pTau.231,
    NEFL = baseline_data$NEFL,
    MAPT = baseline_data$MAPT
    )
)

colours <- c("Baseline" = "#f9a340", "POD 1" = "#7fb6cc")
  
  # Plot
  library(ggplot2)
  p <- ggplot(combined_data, aes(x = factor(timepoint), y = NEFL, color = timepoint)) +
    geom_point(size = 3) +
    geom_line(aes(group = patient_id), color = "grey60") +
    labs(
      title = "NEFL Levels Pre and Post-Operatively",
      x = "Timepoint",
      y = "NEFL",
      color = "Timepoint"
    ) +
    scale_color_manual(values = colours) +
    theme_minimal()
  
  print(p)
  ggsave(filename = "NEFL_Levels_Baseline_Post-op.svg", plot = p, width = 8, height = 6, dpi = 300)

```
## compare biomarkers in the MRI dataset
```{r compare biomarkers}
# Load the MRI clinical data
subject_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/subject_data_IL810_mri.csv")
cog_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/mri_scanner_cog.csv")
# Remove (22) from patient_id in row 21 if present
subject_data$patientid[21] <- gsub("\\(22\\)$", "", subject_data$patientid[21])
cog_data$patientid[21] <- gsub("\\(22\\)$", "", cog_data$patientid[21])

# Merge data based on biomarker and MRI patient IDs -  data structure so repeats for subject for the different timepoints
biomarker_mri_data <- merge(subject_data, combined_data, by.x = "patientid", by.y = "patient_id", all.x = TRUE)
biomarker_cog_mri_data <- merge(cog_data, combined_data, by.x = "patientid", by.y = "patient_id", all.x = TRUE)
#filter for only the preop/baseline data
biomarker_cog_baseline <- merge(cog_data, baseline_data, by.x = "patientid", by.y = "sample_id", all.x = TRUE)
biomarker_cog_POD1 <- merge(cog_data, POD1_data, by.x = "patientid", by.y = "sample_id", all.x = TRUE)
```

```{r}
setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/lm_model/')

baseline_data <- biomarker_cog_mri_data[biomarker_cog_mri_data$timepoint == "Baseline", ]
#load in FC data for ACh
fc_nbm_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/avg_nbm_fc_400_all_subs.csv", header = FALSE)
# load in only sig difference - 

#sig_diff_nbmfc <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/sig_diff_nbm_fc.csv", header = FALSE)
sig_diff_nbmfc <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/nbm_combined_fc_sig_plot.csv", header = FALSE)
## Filtering the FC for only significant ROIs
# If sig_diff_nbmfc is a vector or single-row data frame
sig_fc_nbm <- fc_nbm_data
zero_cols <- as.numeric(sig_diff_nbmfc[1:400, 1]) == 0
# Set those columns to zero across all rows
sig_fc_nbm[, zero_cols] <- 0
roi_names <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/code/schaef_400roi_names.csv", header = FALSE)
colnames(sig_fc_nbm) <- roi_names$V1[1:400]
```
## Run the logistic regression model
```{r run logistic regression}
# Remove columns that are all zeros - only run on significant ROIs
#data_fc <- sig_fc_nbm[, colSums(sig_fc_nbm != 0) > 0]
colnames(fc_nbm_data) <- roi_names$V1[1:400]
#run the glm model
#biomarker_cog_baseline$tmtb
response_var <- "bin_delirium"  # Ensure this matches the column name in your data
#covariates <- "pTau.181 + tmtb"
covariates <- c("pTau.181","patient_age","patient_gender")
glm_nbm_ptau181_preop <- run_glm_model(biomarker_cog_baseline, fc_nbm_data, response_var, covariates)

View(glm_nbm_ptau181_preop$coef_pval_df)
# Extract the interactions between ROI and covariates
interactions <- glm_nbm_ptau181_preop$coef_pval_df[grep(":", glm_nbm_ptau181_preop$coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)
# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)
```



# plot on the brain
```{r plot on the brain}
# Just plot all zeros except for significant interactions
roi_df <- data.frame(
  label = roi_names$V1,
  SignificantCoefficient = 0
)
# Fill in significant coefficients for significant ROIs
roi_df$SignificantCoefficient[roi_df$label %in% sig_interactions$label] <-
  sig_interactions$SignificantCoefficient[match(roi_df$label, sig_interactions$label, nomatch = 0)]

# Get the significant odds ratio
plot_brain_data(roi_df, "PreOp_pTau181_moca_FC_nbM_Sig")
```
# run across all the FC ROIs regardless of SignificantCoefficient
```{r run across all the FC ROIs}
colnames(fc_nbm_data) <- roi_names$V1[1:400]

response_var <- "bin_delirium"  # Ensure this matches the column name in your data
covariates <- c("pTau.217", "patient_age", "patient_gender")
glm_nbm_ptau181_preop <- run_glm_model(biomarker_cog_baseline, fc_nbm_data, response_var, covariates)

View(glm_nbm_ptau181_preop$coef_pval_df)
# Extract the interactions between ROI and covariates
interactions <- glm_nbm_ptau181_preop$coef_pval_df[grep(":", glm_nbm_ptau181_preop$coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)
# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)
#plot significant ones onto brain
data_plot <- data.frame(label = interactions$label, SignificantCoefficient = interactions$SignificantCoefficient)
plot_brain_data(data_plot, "PreOp_pTau217_delirium_FC_nbM_all_age_gender_covariates")


sig_interactions$FDR <- p.adjust(sig_interactions$PValue, method = "fdr")
```
# Run linear model on avg nbM FC and biomarkers
```{r}
# linear model
#avg_fc <- rowMeans(fc_nbm_data, na.rm = TRUE)
# linear model between avg_fc nbm and nefl (and other blood markers)
# Make sure NEFL, patient_age, and patient_gender are in biomarker_cog_baseline
lm_df <- data.frame(
  avg_fc = avg_fc,
  NEFL= biomarker_cog_baseline$NEFL,
  MAPT = biomarker_cog_baseline$MAPT,
  pTau231 = biomarker_cog_baseline$pTau.231,
  pTau181 = biomarker_cog_baseline$pTau.181,
  age = biomarker_cog_baseline$patient_age,
  gender = as.factor(biomarker_cog_baseline$patient_gender)
)
# Run the linear regression
model1 <- lm(avg_fc ~  NEFL + age + gender, data = lm_df)
model2 <- lm(avg_fc ~  MAPT + age + gender, data = lm_df)
model3 <- lm(avg_fc ~  pTau231 + age + gender, data = lm_df)
model4 <- lm(avg_fc ~  pTau181 + age + gender, data = lm_df)
model5 <- lm(avg_fc ~  pTau181*pTau231 + age + gender, data = lm_df)
# Calculate AIC
# Calculate AIC for each
aic1 <- AIC(model1)
aic2 <- AIC(model2)
aic3 <- AIC(model3)
aic4 <- AIC(model4)
aic5 <- AIC(model5)
print(aic5)
# Compare
aic_values <- c(model1 = aic1, model2 = aic2, model3 = aic3, model4 = aic4)
print(aic_values)
# The model with the **lowest AIC** is preferred.
best_model_name <- names(which.min(aic_values))
cat("Best model by AIC:", best_model_name, "\n")

# View the summary
#summary(lm_avgfc_biomarker)
# Print results
print(aic_value)
print(bic_value)


plot_avg_fc_vs_var(avg_fc, biomarker_cog_baseline$pTau.181, biomarker_cog_baseline$bin_delirium, y_label = "Avg FC nbM", plot_title = "Relationship between Avg NBM FC and PreOp pTau181", filename = "avg_nbm_fc_vs_preopptau181.svg")


```
## run lm for biomarkers
```{r}
run_lm_biomarker_fc <- function(
  biomarker_var,           # e.g., "pTau231"
  fc_data,                 # your FC data frame (ROIs as columns)
  biomarker_data,          # your subject data frame
  covariates = c("age", "gender"), # covariates to include
  output_csv = NULL        # optional: filename to save results
) {
  coef_df <- data.frame()
  # Prepare the data frame for modeling
  lm_df <- data.frame(
    age = biomarker_data$patient_age,
    gender = as.factor(biomarker_data$patient_gender)
  )
  # Add the biomarker variable
  lm_df[[biomarker_var]] <- biomarker_data[[biomarker_var]]
  
  for (roi in colnames(fc_data)) {
    lm_df$avg_fc <- fc_data[[roi]]
    formula_str <- paste0("avg_fc ~ ", biomarker_var, " + ", paste(covariates, collapse = " + "))
    lm_fit <- lm(as.formula(formula_str), data = lm_df)
    coef_table <- summary(lm_fit)$coefficients
    # Extract only the biomarker row
    if (biomarker_var %in% rownames(coef_table)) {
      coef_df <- rbind(coef_df, data.frame(
        ROI = roi,
        Estimate = coef_table[biomarker_var, "Estimate"],
        StdError = coef_table[biomarker_var, "Std. Error"],
        tValue = coef_table[biomarker_var, "t value"],
        PValue = coef_table[biomarker_var, "Pr(>|t|)"]
      ))
    }
  }
  # FDR correction
  coef_df$PValue_FDR <- p.adjust(coef_df$PValue, method = "fdr")
  # Optionally save to CSV
  if (!is.null(output_csv)) {
    write.csv(coef_df, output_csv, row.names = FALSE)
  }
  return(coef_df)
}

# Example usage:
# ptau231_results <- run_lm_biomarker_fc(
#   biomarker_var = "pTau231",
#   fc_data = fc_nbm_data,
#   biomarker_data = biomarker_cog_baseline,
#   covariates = c("age", "gender"),
#   output_csv = "LM_eachROI_FCnbM_pTau231_covariates_age_gender.csv"
# )
```


# run the linear model across all ROIs in the FC
```{r}
setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/lm_model/')
# Run the linear regression across all ROIs in the FC
colnames(fc_nbm_data) <- roi_names$V1[1:400]

# Make sure NEFL, patient_age, and patient_gender are in biomarker_cog_baseline
lm_df <- data.frame(
  #fc_nbm_data = fc_nbm_data,
  #NEFL= biomarker_cog_baseline$NEFL,
  #MAPT = biomarker_cog_baseline$MAPT,
  pTau231 = biomarker_cog_baseline$pTau.231,
  age = biomarker_cog_baseline$patient_age,
  gender = as.factor(biomarker_cog_baseline$patient_gender)
)
lm_results <- list()
coef_pval <- list()
all_coef_pval_df <- data.frame()  # Initialize an empty data frame to store all results

# Run the linear regression for each ROI and save outputs
for (roi in colnames(fc_nbm_data)) {
  lm_df$avg_fc <- fc_nbm_data[[roi]]
  lm_fit <- lm(avg_fc ~ NEFL + age + gender, data = lm_df)
  lm_results[[roi]] <- summary(lm_fit)
  coef_table <- summary(lm_fit)$coefficients
  coef_pval[[roi]] <- data.frame(
    ROI = roi,
    Variable = rownames(coef_table),
    Estimate = coef_table[, "Estimate"],
    StdError = coef_table[, "Std. Error"],
    tValue = coef_table[, "t value"],
    PValue = coef_table[, "Pr(>|t|)"]
  )
  all_coef_pval_df <- rbind(all_coef_pval_df, coef_pval[[roi]])
}

# Optionally, save all_coef_pval_df to CSV for later use
# write.csv(all_coef_pval_df, "all_roi_lm_results.csv", row.names = FALSE)
# View the summary
#summary(lm_avgfc_biomarker)
# ...existing code...
nefl_coef_df <- data.frame()  # To store NEFL coefficients for each ROI

for (roi in colnames(fc_nbm_data)) {
  lm_df$avg_fc <- fc_nbm_data[[roi]]
  lm_fit <- lm(avg_fc ~ NEFL + age + gender, data = lm_df)
  coef_table <- summary(lm_fit)$coefficients
  # Extract only the NEFL row
  if ("NEFL" %in% rownames(coef_table)) {
    nefl_coef_df <- rbind(nefl_coef_df, data.frame(
      ROI = roi,
      Estimate = coef_table["NEFL", "Estimate"],
      StdError = coef_table["NEFL", "Std. Error"],
      tValue = coef_table["NEFL", "t value"],
      PValue = coef_table["NEFL", "Pr(>|t|)"]
    ))
  }
}

# Now nefl_coef_df contains the NEFL coefficient for each ROI
View(nefl_coef_df)
# FDR corrected p-values
nefl_coef_df$PValue_FDR <- p.adjust(nefl_coef_df$PValue, method = "fdr")

#find significant relationships
sig_nefl_coef_df <- nefl_coef_df[nefl_coef_df$PValue_FDR < 0.05, ]
write.csv(nefl_coef_df, "LM_eachROI_FCnbM_NEFL_covariates_age_gender.csv", row.names = TRUE)

# run it for MAPT
mapt_coef_df <- data.frame()  # To store NEFL coefficients for each ROI

for (roi in colnames(fc_nbm_data)) {
  lm_df$avg_fc <- fc_nbm_data[[roi]]
  lm_fit <- lm(avg_fc ~ MAPT + age + gender, data = lm_df)
  coef_table <- summary(lm_fit)$coefficients
  # Extract only the NEFL row
  if ("MAPT" %in% rownames(coef_table)) {
    mapt_coef_df <- rbind(mapt_coef_df, data.frame(
      ROI = roi,
      Estimate = coef_table["MAPT", "Estimate"],
      StdError = coef_table["MAPT", "Std. Error"],
      tValue = coef_table["MAPT", "t value"],
      PValue = coef_table["MAPT", "Pr(>|t|)"]
    ))
  }
}

# Now nefl_coef_df contains the NEFL coefficient for each ROI
View(mapt_coef_df)
# FDR corrected p-values
mapt_coef_df$PValue_FDR <- p.adjust(mapt_coef_df$PValue, method = "fdr")

#find significant relationships
sig_mapt_coef_df <- mapt_coef_df[mapt_coef_df$PValue_FDR < 0.05, ]
write.csv(nefl_coef_df, "LM_eachROI_FCnbM_MAPT_covariates_age_gender.csv", row.names = TRUE)

# run it for pTau231
ptau231_coef_df <- data.frame()  # To store NEFL coefficients for each ROI

for (roi in colnames(fc_nbm_data)) {
  lm_df$avg_fc <- fc_nbm_data[[roi]]
  lm_fit <- lm(avg_fc ~ pTau231 + age + gender, data = lm_df)
  coef_table <- summary(lm_fit)$coefficients
  # Extract only the NEFL row
  if ("pTau231" %in% rownames(coef_table)) {
    ptau231_coef_df <- rbind(ptau231_coef_df, data.frame(
      ROI = roi,
      Estimate = coef_table["pTau231", "Estimate"],
      StdError = coef_table["pTau231", "Std. Error"],
      tValue = coef_table["pTau231", "t value"],
      PValue = coef_table["pTau231", "Pr(>|t|)"]
    ))
  }
}
# FDR corrected p-values
ptau231_coef_df$PValue_FDR <- p.adjust(ptau231_coef_df$PValue, method = "fdr")
# Now nefl_coef_df contains the NEFL coefficient for each ROI
View(ptau231_coef_df)


#find significant relationships
sig_ptau231_coef_df <- ptau231_coef_df[ptau231_coef_df$PValue_FDR < 0.05, ]
write.csv(ptau231_coef_df, "LM_eachROI_FCnbM_pTau231_covariates_age_gender.csv", row.names = TRUE)

biomarker_cog_baseline$pTau.181
ptau181_results <- run_lm_biomarker_fc(
  biomarker_var = "pTau.181",
  fc_data = fc_nbm_data,
  biomarker_data = biomarker_cog_baseline,
  covariates = c("age", "gender"),
  output_csv = "LM_eachROI_FCnbM_pTau181_covariates_age_gender.csv")
#significance
sig_ptau181_coef_df <- ptau181_results[ptau181_results$PValue_FDR < 0.05, ]


```

# plot correlation relationship
```{r plot correlation relationship}
plot_avg_fc_vs_var(
  fc_data = fc_nbm_data,                             # your FC matrix/data frame
  y_vector = biomarker_cog_baseline$pTau.181,       # your outcome variable
  group_vector = biomarker_cog_baseline$bin_delirium,   # your group variable (e.g., 0/1)
  y_label = "Pre-Op pTau181",                       # y-axis label
  plot_title = "Relationship between Pre-Op pTau181 and Avg nbM FC", # plot title
  filename = "Avg_fcnbm_vs_preoppTau181.svg"                # output file name
)
```

# just plotting the raw values
```{r plot raw values}
# Plot the raw values of pTau181 pre-op
y_vector <- biomarker_cog_baseline$pTau.181
fc_data <- fc_nbm_data[["17Networks_LH_DorsAttnA_SPL_3"]]
group_vector <- biomarker_cog_baseline$bin_delirium
y_label <- "Pre-Op pTau181"
plot_title <- "Relationship between Pre-Op pTau181 and FC nbM to DorsAttnA_SPL_3"
cor_avg_fc <- cor.test(y_vector, fc_data, use = "complete.obs")
    plot_df <- data.frame(fc_data = fc_data, y_var = y_vector, 
        group = factor(group_vector, levels = c(0, 1), labels = c("Non-Delirious", 
            "Delirious")))
    colours <- c(`Non-Delirious` = "#f9a340", Delirious = "#7fb6cc")
    library(ggplot2)
    p2 <- ggplot(plot_df, aes(x = fc_data, y = y_var, color = group)) + 
        geom_point(size = 3) + geom_smooth(aes(group = 1), method = "lm", 
        se = FALSE, color = "black") + annotate("text", x = max(fc_data, 
        na.rm = TRUE) * 0.8, y = max(plot_df$y_var, na.rm = TRUE) * 
        0.9, label = paste("r =", round(cor_avg_fc$estimate, 
        2), "\np =", signif(cor_avg_fc$p.value, 3)), color = "black", 
        size = 5) + labs(title = plot_title, x = "FC ROI", y = y_label, 
        color = "Delirium Group") + scale_color_manual(values = colours) + 
        theme_minimal()
    print(p2)
    ggsave(filename = "FC nbM to DorsAttnA SPL3 vs Pre-Op pTau181.svg", plot = p2, width = 8, height = 6, dpi = 300)
```

# relationship between pTau181 and 
```{r relationship between pTau181 and MMN}
p1 <-ggplot(biomarker_cog_baseline, aes(x = pTau.181, y = overall_peak_drs_calc2)) +
  geom_point(size = 3, color = "#7fb6cc") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(
    title = "Relationship between Pre-Op pTau181 and Peak DRS",
    x = "Pre-Op pTau181",
    y = "Peak DRS"
  ) +
  theme_minimal()
  print(p1)
  ggsave(filename = "Relationship_peakDrsCacl2_PreOp_pTau181.svg",plot=p1, width = 8, height = 6, dpi = 300)

cor_res <- cor.test(biomarker_cog_POD1$pTau.181, biomarker_cog_POD1$overall_peak_drs_calc2, method = "spearman")

p1 <- ggplot(biomarker_cog_POD1, aes(x = pTau.181, y = overall_peak_drs_calc2)) +
  geom_point(size = 3, color = "#7fb6cc") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  annotate(
    "text",
    x = max(biomarker_cog_POD1$pTau.181, na.rm = TRUE) * 0.8,
    y = max(biomarker_cog_POD1$overall_peak_drs_calc2, na.rm = TRUE) * 0.9,
    label = paste0("rho = ", round(cor_res$estimate, 2), "\np = ", signif(cor_res$p.value, 3)),
    color = "black",
    size = 5
  ) +
  labs(
    title = "Relationship between Post-Op pTau181 and Peak DRS",
    x = "Post-Op pTau181",
    y = "Peak DRS"
  ) +
  theme_minimal()

print(p1)

  ggsave(filename = "Relationship_peakDrsCacl2_PostOp_pTau181.svg",plot=p1, width = 8, height = 6, dpi = 300)


```

# Relationship between pTau181 and Cognition schaef_400roi_names
```{r relationship between pTau181 and cognition}
# Load the MRI clinical data
data_fc <- sig_fc_nbm[, colSums(sig_fc_nbm != 0) > 0]
#run the glm model
#biomarker_cog_baseline$tmtb
response_var <- "tmtb"  # Ensure this matches the column name in your data
covariates <- "pTau.181"
glm_nbm_ptau181_preop_tmtb <- run_linear_reg(biomarker_cog_baseline, data_fc, response_var, covariates)

View(glm_nbm_ptau181_preop_tmtb$coef_pval_df)
# Extract the interactions between ROI and covariates
interactions <- glm_nbm_ptau181_preop_tmtb$coef_pval_df[grep(":", glm_nbm_ptau181_preop_tmtb$coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)
# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)

response_var <- "pTau.181"  # Ensure this matches the column name in your data
covariates <- c("patient_age", "patient_gender")
glm_nbm_ptau181_preop_tmtb <- run_linear_reg(biomarker_cog_baseline, fc_nbm_data, response_var, covariates)

View(glm_nbm_ptau181_preop_tmtb$coef_pval_df)
# Extract the interactions between ROI and covariates
# Extract only the coefficients for the FC ROI variables (not interactions or covariates)
fc_roi_names <- colnames(fc_nbm_data)
# In the coef_pval_df, the Variable column should match the FC ROI names exactly for main effects
fc_estimates <- glm_nbm_ptau181_preop_tmtb$coef_pval_df[
  grepl("^X", glm_nbm_ptau181_preop_tmtb$coef_pval_df$Variable) & !grepl(":", glm_nbm_ptau181_preop_tmtb$coef_pval_df$Variable),
]
View(fc_estimates)


fc_estimates$SignificantCoefficient <- ifelse(fc_estimates$PValue < 0.05, fc_estimates$Estimate, 0)

sig_interactions <- fc_estimates[fc_estimates$PValue < 0.05, ]
View(sig_interactions)
fc_estimates$label <- sub(":.*", "", fc_estimates$Variable)
fc_estimates$label <- sub("X", "", fc_estimates$label)

data_plot <- data.frame(label = fc_estimates$label, SignificantCoefficient = fc_estimates$SignificantCoefficient)
plot_brain_data(data_plot, "PreOp_pTau181_FC_nbM_all_age_gender_covariates")
```
# run correlation across each FC nbM ROI to pTau181
```{r}
ptau_vec <- biomarker_cog_baseline$pTau.181

cor_results <- sapply(fc_nbm_data, function(roi) {
  cor.test(roi, ptau_vec, use = "complete.obs")$estimate
})

pval_results <- sapply(fc_nbm_data, function(roi) {
  cor.test(roi, ptau_vec, use = "complete.obs")$p.value
})

# Combine into a data frame for easy viewing
colnames(fc_nbm_data)<-roi_names$V1[1:400]
cor_summary <- data.frame(
  ROI = colnames(fc_nbm_data),
  Correlation = cor_results,
  PValue = pval_results
)

View(cor_summary)
# filter for significant correlations
sig_cor_summary <- cor_summary[cor_summary$PValue < 0.05, ]
View(sig_cor_summary)

#plot onto the brain
# Create a data frame with all ROI labels and fill with 0
sig_corr_data <- data.frame(
  label = roi_names$V1[1:400],
  SignificantCoefficient = 0
)
# Fill in significant correlations for ROIs present in sig_cor_summary
match_idx <- match(sig_cor_summary$ROI, sig_corr_data$label)
sig_corr_data$SignificantCoefficient[match_idx[!is.na(match_idx)]] <- sig_cor_summary$Correlation[!is.na(match_idx)]

#plot significant ones onto brain
plot_brain_data(sig_corr_data, "Correlation_Sig_PreOp_pTau181_FC_nbM_allROIs")


# FDR correction on correlation p-values
cor_summary$PValue_FDR <- p.adjust(cor_summary$PValue, method = "fdr")

# Filter for significant correlations after FDR correction
sig_cor_summary_fdr <- cor_summary[cor_summary$PValue_FDR < 0.05, ]
View(sig_cor_summary_fdr)

# Update the plotting data for FDR-significant ROIs
sig_corr_data_fdr <- data.frame(
  label = roi_names$V1[1:400],
  SignificantCoefficient = 0
)
match_idx_fdr <- match(sig_cor_summary_fdr$ROI, sig_corr_data_fdr$label)
sig_corr_data_fdr$SignificantCoefficient[match_idx_fdr[!is.na(match_idx_fdr)]] <- sig_cor_summary_fdr$Correlation[!is.na(match_idx_fdr)]

# Plot FDR-significant correlations onto the brain
plot_brain_data(sig_corr_data_fdr, "Correlation_FDR_Sig_PreOp_pTau181_FC_nbM_allROIs")


```
# Relationship between NEFL biomarker and avg. FC
```{r}
avg_nbm_fc <- rowMeans(fc_nbm_data, na.rm = TRUE)
#run the glm model
#biomarker_cog_baseline$tmtb
response_var <- "avg_nbm_fc"  # Ensure this matches the column name in your data
covariates <- c("patient_age", "patient_gender")
glm_nbm_ptau181_preop_tmtb <- run_linear_reg(biomarker_cog_baseline, fc_nbm_data, response_var, covariates)

View(glm_nbm_ptau181_preop_tmtb$coef_pval_df)
# Extract the interactions between ROI and covariates
# Extract only the coefficients for the FC ROI variables (not interactions or covariates)
fc_roi_names <- colnames(fc_nbm_data)
# In the coef_pval_df, the Variable column should match the FC ROI names exactly for main effects
fc_estimates <- glm_nbm_ptau181_preop_tmtb$coef_pval_df[
  grepl("^X", glm_nbm_ptau181_preop_tmtb$coef_pval_df$Variable) & !grepl(":", glm_nbm_ptau181_preop_tmtb$coef_pval_df$Variable),
]
View(fc_estimates)


fc_estimates$SignificantCoefficient <- ifelse(fc_estimates$PValue < 0.05, fc_estimates$Estimate, 0)

sig_interactions <- fc_estimates[fc_estimates$PValue < 0.05, ]
View(sig_interactions)
fc_estimates$label <- sub(":.*", "", fc_estimates$Variable)
fc_estimates$label <- sub("X", "", fc_estimates$label)

data_plot <- data.frame(label = fc_estimates$label, SignificantCoefficient = fc_estimates$SignificantCoefficient)
plot_brain_data(data_plot, "PreOp_pTau181_FC_nbM_all_age_gender_covariates")

```
# Relationship between pTau181 and PC
```{r relationship between pTau181 and PC}
## Load in the data
pc_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/avg_pc_all_subs.csv", header = FALSE)
pc_data <- pc_data[,1:400]
colnames(pc_data) <- roi_names$V1[1:400]
biomarker_cog_baseline <- biomarker_cog_baseline[-88,] #need to remove sub-164 as it is not in the pc data
biomarker_cog_POD1 <- biomarker_cog_POD1[-88,] #need to remove sub-164 as it is not in the pc data
#filter fc_data to match mmn_data
#whole brain avg. pc for each subject

ptau_vec <- biomarker_cog_baseline$pTau.181

cor_results <- sapply(pc_data, function(roi) {
  cor.test(roi, ptau_vec, use = "complete.obs")$estimate
})

pval_results <- sapply(pc_data, function(roi) {
  cor.test(roi, ptau_vec, use = "complete.obs")$p.value
})

# Combine into a data frame for easy viewing
cor_summary <- data.frame(
  ROI = colnames(pc_data),
  Correlation = cor_results,
  PValue = pval_results
)

View(cor_summary)
# filter for significant correlations
sig_cor_summary <- cor_summary[cor_summary$PValue < 0.05, ]
View(sig_cor_summary)

#plot onto the brain
# Create a data frame with all ROI labels and fill with 0
#sig_corr_data <- data.frame(
  #label = roi_names$V1[1:400],
  #SignificantCoefficient = 0
#)
# Fill in significant correlations for ROIs present in sig_cor_summary
#match_idx <- match(sig_cor_summary$ROI, sig_corr_data$label)
#sig_corr_data$SignificantCoefficient[match_idx[!is.na(match_idx)]] <- sig_cor_summary$Correlation[!is.na(match_idx)]

#plot significant ones onto brain
#plot_brain_data(sig_corr_data, "Correlation_Sig_PreOp_pTau181_FC_nbM_allROIs")


# FDR correction on correlation p-values
cor_summary$PValue_FDR <- p.adjust(cor_summary$PValue, method = "fdr")

# Filter for significant correlations after FDR correction
sig_cor_summary_fdr <- cor_summary[cor_summary$PValue_FDR < 0.05, ]
View(sig_cor_summary_fdr)

# Update the plotting data for FDR-significant ROIs
sig_corr_data_fdr <- data.frame(
  label = roi_names$V1[1:400],
  SignificantCoefficient = 0
)
match_idx_fdr <- match(sig_cor_summary_fdr$ROI, sig_corr_data_fdr$label)
sig_corr_data_fdr$SignificantCoefficient[match_idx_fdr[!is.na(match_idx_fdr)]] <- sig_cor_summary_fdr$Correlation[!is.na(match_idx_fdr)]

# Plot FDR-significant correlations onto the brain
plot_brain_data(sig_corr_data_fdr, "Correlation_FDR_Sig_PreOp_pTau181_avgPC_overtime")

```

# Run linear model on avg PC and biomarkers
```{r}
# linear model
#avg_fc <- rowMeans(fc_nbm_data, na.rm = TRUE)
# linear model between avg_fc nbm and nefl (and other blood markers)
# Make sure NEFL, patient_age, and patient_gender are in biomarker_cog_baseline
avg_pc <- rowMeans(pc_data, na.rm = TRUE)

lm_df <- data.frame(
  avg_pc = avg_pc,
  NEFL= biomarker_cog_baseline$NEFL,
  MAPT = biomarker_cog_baseline$MAPT,
  pTau231 = biomarker_cog_baseline$pTau.231,
  pTau181 = biomarker_cog_baseline$pTau.181,
  age = biomarker_cog_baseline$patient_age,
  gender = as.factor(biomarker_cog_baseline$patient_gender)
)
# Run the linear regression
model1 <- lm(avg_pc ~  NEFL + age + gender, data = lm_df)
model2 <- lm(avg_pc ~  MAPT + age + gender, data = lm_df)
model3 <- lm(avg_pc ~  pTau231 + age + gender, data = lm_df)
model4 <- lm(avg_pc ~  pTau181 + age + gender, data = lm_df)
model5 <- lm(avg_pc ~  pTau181*pTau231 + age + gender, data = lm_df)
# Calculate AIC
# Calculate AIC for each
aic1 <- AIC(model1)
aic2 <- AIC(model2)
aic3 <- AIC(model3)
aic4 <- AIC(model4)
aic5 <- AIC(model5)
print(aic5)
# Compare
aic_values <- c(model1 = aic1, model2 = aic2, model3 = aic3, model4 = aic4, model5= aic5)
print(aic_values)
# The model with the **lowest AIC** is preferred.
best_model_name <- names(which.min(aic_values))
cat("Best model by AIC:", best_model_name, "\n")

# View the summary
#summary(lm_avgfc_biomarker)
# Print results
print(aic_value)
print(bic_value)


plot_avg_fc_vs_var(avg_fc, biomarker_cog_baseline$pTau.181, biomarker_cog_baseline$bin_delirium, y_label = "Avg FC nbM", plot_title = "Relationship between Avg NBM FC and PreOp pTau181", filename = "avg_nbm_fc_vs_preopptau181.svg")

model1$Pval_FDR <- p.adjust(summary(model1)$coefficients[,4], method = "fdr")
```