---
title: "Logistic Regression Modelling for Pre-Op Delirium"
---

```{r load libraries}
#library(lmPerm)
library(broom)
library(ggplot2)
library(ggsegSchaefer)
```
## Plot on the Brain Significant
```{r}
# Plot the significant coefficients on the brain
plot_brain_data <- function(data, filename) {
  library(ggplot2)
  library(ggseg)
  library(ggsegSchaefer)
  library(dplyr)
  library(stringr)
  
  # Ensure the data has a 'label' column
  if (!"label" %in% colnames(data)) {
    stop("The data must contain a 'label' column.")
  }
  
  # Prepare the data for plotting
  data_plot <- data %>%
    mutate(label = ifelse(str_detect(label, "_LH"), paste0("lh_", label), paste0("rh_", label)))
  
  # Print the data frame to verify
  View(data_plot)
  
  # Create the plot
  p1 <- ggplot(data_plot, aes(fill = SignificantCoefficient)) +
    geom_brain(atlas = schaefer17_400, show.legend = TRUE) +
    #scale_fill_gradient2(low = "#CA1C6A", mid = "white", high = "#173F6C", midpoint = 0) + #plotting the pink - blue colour scales
    scale_fill_gradient2(low = "#C03B2B", mid = "white", high = "#0C7D79", midpoint = 0) + #plotting the red - green colour scales
    theme_void() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 14),
          legend.key.size = unit(1.5, "cm"))+
          labs(fill = "Beta Coefficient")  # Set the legend title
  
  # Print the plot
    print(p1)
    setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Manuscripts/AAS_Manuscript/Figures/Supplementary_Figs')
   #setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
  ggsave(paste0(filename, ".svg"), plot = p1, width = 10, height = 10)
}

plot_brain_data_oddsratio <- function(data, filename) {
  library(ggplot2)
  library(ggseg)
  library(ggsegSchaefer)
  library(dplyr)
  library(stringr)
  
  # Ensure the data has a 'label' column
  if (!"label" %in% colnames(data)) {
    stop("The data must contain a 'label' column.")
  }
  
  # Prepare the data for plotting
  data_plot <- data[1:400,] %>%
    mutate(label = ifelse(str_detect(label, "_LH"), paste0("lh_", label), paste0("rh_", label)))
  
  # Print the data frame to verify
  View(data_plot)

 # Create the plot
p2 <- ggplot(data_plot, aes(fill = SignificantCoefficient)) +
    geom_brain(atlas = schaefer17_400, show.legend = TRUE) +
    scale_fill_gradientn(
      colors = c("#C03B2B", "white", "#0C7D79"),  # Red, White, Green
      values = c(0.9, 0, 1),  # Map 0 to white, <1 to red, >1 to green
      limits = c(min(data_plot$SignificantCoefficient, na.rm = TRUE), 
                max(data_plot$SignificantCoefficient, na.rm = TRUE))  # Set limits
    ) +
    theme_void() +
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      legend.key.size = unit(1.5, "cm")
    ) +
    labs(fill = "Log Odds Ratio")  # Set the legend title

  # Print the plot
  print(p2)
  #this is for saving the supplementary Figs.
  setwd('/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Manuscripts/AAS_Manuscript/Figures/Supplementary_Figs')
  #setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
  ggsave(paste0(filename, ".svg"), plot = p2, width = 10, height = 10)
}
```

## Function Code for Logistic Regression
```{r}
## Define the function
run_glm_model <- function(subject_data, data_fc, response_var, covariates) {
  # Ensure variable names are valid
  names(subject_data) <- make.names(names(subject_data))
  names(data_fc) <- make.names(names(data_fc))
  response_var <- make.names(response_var)
  covariates <- make.names(covariates)
  
  # Extract the response variable
  response <- subject_data[[response_var]]
  
  # Combine the data
  data_comb <- cbind(response, data_fc, subject_data[covariates])
  data_model_df <- data.frame(data_comb)
  # Define predictor variables
  predictor_vars <- names(data_comb)[names(data_comb) != "response"]

  # Define the function to run the logistic regression model with permutation
  run_glm <- function(var, data_comb) {
    formula <- as.formula(paste("response ~", var, "*", paste(covariates, collapse = " + ")))
    model <- glm(formula, data = data_comb, family = binomial(link = "logit"))
    return(model)
  }
  
  results <- lapply(predictor_vars, function(var) {
    glm <- run_glm(var, data_comb)
    summary_model <- summary(glm)
    marginal_effects <- predict(glm, data_comb,type = "response")
    return(list(model = glm, summary = summary_model, marginal_effects = marginal_effects))
  })
  names(results) <- predictor_vars  # Assign names to the results list
  # Extract beta coefficients and p-values
  coef_pval_list <- lapply(results, function(res) {
    coef_table <- summary(res$model)$coefficients
    data.frame(
      Variable = rownames(coef_table),
      Beta = coef_table[, "Estimate"],
      PValue = coef_table[, "Pr(>|z|)"],
      OddsRatio = exp(coef_table[, "Estimate"]) # Calculate odds ratio
    )
  })
  #Extract AIC Values
    aic_list <- lapply(results, function(res) {
      aic <- AIC(res$model)
      return(aic)
    })
  names(aic_list) <- predictor_vars  # Assign names to the AIC list
  # Combine all results into a single data frame
  coef_pval_df <- do.call(rbind, coef_pval_list)
  
  return(list(results = results, coef_pval_df = coef_pval_df,data_model_df = data_model_df, aic_list = aic_list))
}
```
## Simple Logistic Regression Model - loops through regions without covariates
```{r}
# run logistic regression for just LC FC
# Assuming subject_data, data_fc, response_var, and covariates are defined
## Define the function
run_glm_simple_model <- function(subject_data, data_fc, response_var) {
  # Ensure variable names are valid
  names(subject_data) <- make.names(names(subject_data))
  names(data_fc) <- make.names(names(data_fc))
  response_var <- make.names(response_var)
  
  # Extract the response variable
  response <- subject_data[[response_var]]
  
  # Combine the data
  data_comb <- cbind(response, data_fc)
  data_model_df <- data.frame(data_comb)
  # Define predictor variables
  predictor_vars <- names(data_comb)[names(data_comb) != "response"]

  # Define the function to run the logistic regression model with permutation
  run_glm <- function(var, data_comb) {
    formula <- as.formula(paste("response ~", var))
    model <- glm(formula, data = data_comb, family = binomial(link = "logit"))
    return(model)
  }
  
  results <- lapply(predictor_vars, function(var) {
    glm <- run_glm(var, data_comb)
    summary_model <- summary(glm)
    marginal_effects <- predict(glm, data_comb, type = "response")
    return(list(model = glm, summary = summary_model, marginal_effects = marginal_effects)) # nolint
  })
  names(results) <- predictor_vars  # Assign names to the results list
  # Extract beta coefficients and p-values
  coef_pval_list <- lapply(results, function(res) {
    coef_table <- summary(res$model)$coefficients
    data.frame(
      Variable = rownames(coef_table),
      Beta = coef_table[, "Estimate"],
      PValue = coef_table[, "Pr(>|z|)"],
      OddsRatio = exp(coef_table[, "Estimate"]) # Calculate odds ratio
    )
  })
  
  # Combine all results into a single data frame
  coef_pval_df <- do.call(rbind, coef_pval_list)
  #print of AIC for each variable too
  aic_list <- lapply(results, function(res) {
    aic <- AIC(res$model)
    return(aic)
  })

  return(list(results = results, coef_pval_df = coef_pval_df,data_model_df = data_model_df, aic_list = aic_list))
}
```
## Logistic Regression Model with Covariate as + not an interaction term
```{r}
## Define the function
run_glm_model2 <- function(subject_data, data_fc, response_var, covariates) {
  # Ensure variable names are valid
  names(subject_data) <- make.names(names(subject_data))
  names(data_fc) <- make.names(names(data_fc))
  response_var <- make.names(response_var)
  covariates <- make.names(covariates)
  
  # Extract the response variable
  response <- subject_data[[response_var]]
  
  # Combine the data
  data_comb <- cbind(response, data_fc, subject_data[covariates])
  data_model_df <- data.frame(data_comb)
  # Define predictor variables
  predictor_vars <- names(data_comb)[names(data_comb) != "response"]

  # Define the function to run the logistic regression model with permutation
  run_glm <- function(var, data_comb) {
    formula <- as.formula(paste("response ~", var, "+", paste(covariates, collapse = " + ")))
    model <- glm(formula, data = data_comb, family = binomial(link = "logit"))
    return(model)
  }
  
  results <- lapply(predictor_vars, function(var) {
    glm <- run_glm(var, data_comb)
    summary_model <- summary(glm)
    marginal_effects <- predict(glm, data_comb,type = "response")
    return(list(model = glm, summary = summary_model, marginal_effects = marginal_effects))
  })
  names(results) <- predictor_vars  # Assign names to the results list
  # Extract beta coefficients and p-values
  coef_pval_list <- lapply(results, function(res) {
    coef_table <- summary(res$model)$coefficients
    data.frame(
      Variable = rownames(coef_table),
      Beta = coef_table[, "Estimate"],
      PValue = coef_table[, "Pr(>|z|)"],
      OddsRatio = exp(coef_table[, "Estimate"]) # Calculate odds ratio
    )
  })
  #Extract AIC Values
    aic_list <- lapply(results, function(res) {
      aic <- AIC(res$model)
      return(aic)
    })
  names(aic_list) <- predictor_vars  # Assign names to the AIC list
  # Combine all results into a single data frame
  coef_pval_df <- do.call(rbind, coef_pval_list)
  
  return(list(results = results, coef_pval_df = coef_pval_df,data_model_df = data_model_df, aic_list = aic_list))
}
```
## Load the data and covariates
```{r load_data}
## Load in the data
#pc_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/avg_pc_all_subs.csv", header = FALSE)
pc_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/csf_regress_redo/avg_pc_all_subs.csv", header = FALSE)
#whole brain avg. pc for each subject
avg_pc <- rowMeans(pc_data, na.rm = TRUE)
# Convert avg_pc to a data frame
avg_pc <- data.frame(avg_pc)
## Clinical Data
subject_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/subject_data_IL810_mri.csv", header=TRUE)
subject_data <- subject_data[-88,] #need to remove sub-164 as it is not in the pc data
# cog data 
cog_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/mri_scanner_cog.csv")
cog_data <- cog_data[-88,]
#load in the ROI names
roi_names <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/code/schaef_400roi_names.csv", header = FALSE)
#rename colum names in pc_data with roi_names
colnames(pc_data) <- roi_names$V1
## oz delta data
oz_delta_orig <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/Oz_deltapower.csv")
# Remove "_" in subject names
oz_delta_orig$Sub <- gsub("_", "", oz_delta_orig$Sub)
# Filter for MRI subjects
oz_delta <- oz_delta_orig[oz_delta_orig$Sub %in% subject_data$patientid, ]
# Merge oz_delta with subject_data based on patientid
subject_data_ozdelta <- merge(subject_data, oz_delta, by.x = "patientid", by.y = "Sub", all.x = TRUE)
```

## Run Log Reg for each ROI (PC) * Diff_OzDeltaPow
```{r}
# Assuming subject_data, data_fc, response_var, and covariates are defined
covariates <- c("Diff_OzDeltaPow")
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_ozdelta_pc <- run_glm_model(subject_data_ozdelta, pc_data, response_var, covariates) # nolint
data_model_df = glm_model_ozdelta_pc$data_model_df #that's the actual data that went into the model
View(glm_model_ozdelta_pc$coef_pval_df)

# Extract the interactions between ROI and covariates
interactions <- glm_model_ozdelta_pc$coef_pval_df[grep(":", glm_model_ozdelta_pc$coef_pval_df$Variable), ]
View(interactions)
# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)
# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)
# FDR correction for multiple comparisons
interactions$FDR_PValue <- p.adjust(interactions$PValue, method = "fdr")
# Add a column to indicate FDR-significant coefficients
interactions$SignificantCoefficient_FDR <- ifelse(interactions$FDR_PValue < 0.05, interactions$Beta, 0)

# Extract significant interactions after FDR correction
sig_interactions <- interactions[interactions$FDR_PValue < 0.05, ]
View(sig_interactions)
```
```{r}
## Plotting the figures
# Get the significant odds ratio
interactions$sig_OddsRatio <- ifelse(interactions$PValue < 0.05, interactions$OddsRatio, 0)
data <- data.frame(label = interactions$label, SignificantCoefficient = interactions$sig_OddsRatio)

plot_brain_data_oddsratio(data, "PC_ROI_ozdeltapower_sigodds")

data2 <- data.frame(label = interactions$label, SignificantCoefficient = interactions$SignificantCoefficient)

plot_brain_data(data2, "PC_ROI_ozdeltapower_sigcoeffs")
# Use the actual data that went into the model, ensuring variable names match
```

## Need to run model fit for each variable to predict delirium outcome
Iterate through each variable & determine the best model for fit- comparison with AIC
```{r}
response_var <- "bin_delirium"
# Model 1: bin delirium ~ PC only
model1 <- run_glm_simple_model(subject_data,pc_data,response_var)
# Model 2: bin delirium ~ Diff_OzDeltaPow only
model2 <- glm(bin_delirium ~ Diff_OzDeltaPow, data=subject_data_ozdelta,family=binomial(link="logit")) # nolint
# model with the interaction
covariates <- c("Diff_OzDeltaPow")
# Model 3: bin delirium ~ PC * Diff_OzDeltaPow
model3 <- run_glm_model(subject_data_ozdelta, pc_data, response_var, covariates) # nolint
# Model 4: bin delirium ~ PC + Diff_OzDeltaPow
model4 <- run_glm_model2(subject_data_ozdelta, pc_data, response_var, covariates) # nolint


# Extract AIC values for each predictor in model1
aic_model1 <- sapply(model1$results, function(res) AIC(res$model))
# Extract AIC values for each predictor in model3
aic_model3 <- sapply(model3$results, function(res) AIC(res$model))
aic_model4 <- sapply(model4$results, function(res) AIC(res$model))
# Extract AIC value for model2
aic_model2 <- AIC(model2)

# Combine AIC values into a data frame
# Combine all predictors and models into a single vector of equal length
predictor_names <- c(names(aic_model1), "Diff_OzDeltaPow", names(aic_model3), names(aic_model4))
model_names <- c(
  rep("Model 1", length(aic_model1)),
  "Model 2",
  rep("Model 3", length(aic_model3)),
  rep("Model 4", length(aic_model4))
)
aic_values <- c(aic_model1, aic_model2, aic_model3, aic_model4)

# Ensure all vectors are the same length
stopifnot(length(predictor_names) == length(model_names), length(model_names) == length(aic_values))

aic_comparison <- data.frame(
  Predictor = predictor_names,
  Model = model_names,
  AIC = aic_values
)



    # Identify the best model based on the lowest AIC
    best_model <- aic_comparison[which.min(aic_comparison$AIC), ]

    # Print the best model
    print(paste("The best model for", best_model$Model ,"is for variable:", best_model$Predictor, "with AIC:", best_model$AIC))


        # Plot the AIC comparisons with color based on the model
        library(ggplot2)
        aic_plot <-ggplot(aic_comparison, aes(x = reorder(Predictor,AIC), y = AIC, color = Model)) +
          geom_point(size = 3) +
          coord_flip() +
          labs(title = "AIC Comparison Across Models",
               x = "Predictor",
               y = "AIC",
               color = "Model") +
          theme_minimal() +
          theme(axis.text.x = element_text(angle = 90, hjust = 1))
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
  filename="AIC_comparison_PC_ROI_delirium_severity"
  ggsave(paste0(filename, ".png"),plot=aic_plot, width = 10, height = 10)

# Filter predictors in model 3 where AIC is less than model 2
predictors_better_than_model2 <- names(aic_model3[aic_model3 < aic_model2])

# Print the predictors
print(predictors_better_than_model2)
```
# AIC model summary for all models
```{r}
# =============================================================================
# COMPREHENSIVE AIC ANALYSIS FOR ALL MODELS
# =============================================================================

# Extract AIC values for all models
# Model 1: PC only
aic_model1 <- sapply(model1$results, function(res) AIC(res$model))

# Model 2: OzDelta only  
aic_model2 <- AIC(model2)

# Model 3: PC * OzDelta (additive)
aic_model3 <- sapply(model3$results, function(res) AIC(res$model))

# Model 4: PC + OzDelta (interaction)
aic_model4 <- sapply(model4$results, function(res) AIC(res$model))

# =============================================================================
# INDIVIDUAL MODEL SUMMARIES
# =============================================================================

cat("=== MODEL 1 (PC ONLY) SUMMARY ===\n")
model1_summary <- list(
  n_models = length(aic_model1),
  best_aic = min(aic_model1),
  worst_aic = max(aic_model1),
  mean_aic = mean(aic_model1),
  median_aic = median(aic_model1),
  sd_aic = sd(aic_model1),
  best_model = names(which.min(aic_model1))
)
print(model1_summary)

cat("\n=== MODEL 2 (OZDELTA ONLY) SUMMARY ===\n")
cat("AIC:", aic_model2, "\n")

cat("\n=== MODEL 3 (PC + OZDELTA ADDITIVE) SUMMARY ===\n")
model3_summary <- list(
  n_models = length(aic_model3),
  best_aic = min(aic_model3),
  worst_aic = max(aic_model3),
  mean_aic = mean(aic_model3),
  median_aic = median(aic_model3),
  sd_aic = sd(aic_model3),
  best_model = names(which.min(aic_model3))
)
print(model3_summary)

cat("\n=== MODEL 4 (PC * OZDELTA INTERACTION) SUMMARY ===\n")
model4_summary <- list(
  n_models = length(aic_model4),
  best_aic = min(aic_model4),
  worst_aic = max(aic_model4),
  mean_aic = mean(aic_model4),
  median_aic = median(aic_model4),
  sd_aic = sd(aic_model4),
  best_model = names(which.min(aic_model4))
)
print(model4_summary)

# =============================================================================
# DETAILED MODEL COMPARISONS
# =============================================================================

# Create comprehensive comparison table
create_model_comparison <- function() {
  # Get representative AIC for each model type
  model1_best <- min(aic_model1)
  model1_best_name <- names(which.min(aic_model1))
  
  model3_best <- min(aic_model3)
  model3_best_name <- names(which.min(aic_model3))
  
  model4_best <- min(aic_model4)
  model4_best_name <- names(which.min(aic_model4))
  
  # Create comparison data frame
  comparison_df <- data.frame(
    Model_Type = c("Model 1 (PC only)", 
                   "Model 2 (OzDelta only)", 
                   "Model 3 (PC + OzDelta)", 
                   "Model 4 (PC * OzDelta)"),
    Best_Predictor = c(model1_best_name, 
                       "Diff_OzDeltaPow", 
                       model3_best_name, 
                       model4_best_name),
    Best_AIC = c(model1_best, 
                 aic_model2, 
                 model3_best, 
                 model4_best),
    N_Models_Tested = c(length(aic_model1), 
                        1, 
                        length(aic_model3), 
                        length(aic_model4)),
    stringsAsFactors = FALSE
  )
  
  # Add delta AIC and rankings
  comparison_df$Delta_AIC <- comparison_df$Best_AIC - min(comparison_df$Best_AIC)
  comparison_df$Rank <- rank(comparison_df$Best_AIC)
  
  # Add AIC weights
  comparison_df$AIC_Weight <- exp(-0.5 * comparison_df$Delta_AIC) / 
                             sum(exp(-0.5 * comparison_df$Delta_AIC))
  
  # Sort by AIC
  comparison_df <- comparison_df[order(comparison_df$Best_AIC), ]
  
  return(comparison_df)
}

cat("\n=== OVERALL MODEL COMPARISON ===\n")
overall_comparison <- create_model_comparison()
print(overall_comparison)

# =============================================================================
# STATISTICAL SIGNIFICANCE OF DIFFERENCES
# =============================================================================

cat("\n=== MODEL COMPARISON ANALYSIS ===\n")

# Find overall best model
best_overall <- overall_comparison[1, ]
cat("BEST OVERALL MODEL:\n")
cat("Type:", best_overall$Model_Type, "\n")
cat("Predictor:", best_overall$Best_Predictor, "\n")
cat("AIC:", round(best_overall$Best_AIC, 2), "\n")

# Models with substantial support (Delta AIC < 2)
competitive_models <- overall_comparison[overall_comparison$Delta_AIC < 2, ]
cat("\nCOMPETITIVE MODELS (ΔAICc < 2):\n")
if(nrow(competitive_models) > 1) {
  print(competitive_models[, c("Model_Type", "Best_Predictor", "Best_AIC", "Delta_AIC")])
} else {
  cat("Only one model has substantial support\n")
}

# Models with some support (Delta AIC < 7)
some_support <- overall_comparison[overall_comparison$Delta_AIC < 7, ]
cat("\nMODELS WITH SOME SUPPORT (ΔAICc < 7):\n")
print(some_support[, c("Model_Type", "Best_Predictor", "Best_AIC", "Delta_AIC")])

# =============================================================================
# DETAILED BREAKDOWN BY MODEL TYPE
# =============================================================================

cat("\n=== DETAILED BREAKDOWN ===\n")

# Top 3 models from each category
cat("TOP 3 MODELS FROM EACH CATEGORY:\n\n")

# Model 1 top 3
cat("Model 1 (PC only) - Top 3:\n")
model1_top3 <- head(sort(aic_model1), 3)
for(i in 1:length(model1_top3)) {
  cat(sprintf("%d. %s: AIC = %.2f\n", i, names(model1_top3)[i], model1_top3[i]))
}

cat("\nModel 2 (OzDelta only):\n")
cat(sprintf("1. Diff_OzDeltaPow: AIC = %.2f\n", aic_model2))

# Model 3 top 3
cat("\nModel 3 (PC + OzDelta) - Top 3:\n")
model3_top3 <- head(sort(aic_model3), 3)
for(i in 1:length(model3_top3)) {
  cat(sprintf("%d. %s: AIC = %.2f\n", i, names(model3_top3)[i], model3_top3[i]))
}

# Model 4 top 3
cat("\nModel 4 (PC * OzDelta) - Top 3:\n")
model4_top3 <- head(sort(aic_model4), 3)
for(i in 1:length(model4_top3)) {
  cat(sprintf("%d. %s: AIC = %.2f\n", i, names(model4_top3)[i], model4_top3[i]))
}

# =============================================================================
# SUMMARY STATISTICS TABLE
# =============================================================================

cat("\n=== SUMMARY STATISTICS BY MODEL TYPE ===\n")
summary_stats <- data.frame(
  Model_Type = c("Model 1", "Model 2", "Model 3", "Model 4"),
  N_Models = c(length(aic_model1), 1, length(aic_model3), length(aic_model4)),
  Best_AIC = c(min(aic_model1), aic_model2, min(aic_model3), min(aic_model4)),
  Mean_AIC = c(mean(aic_model1), aic_model2, mean(aic_model3), mean(aic_model4)),
  SD_AIC = c(sd(aic_model1), 0, sd(aic_model3), sd(aic_model4)),
  Range_AIC = c(diff(range(aic_model1)), 0, diff(range(aic_model3)), diff(range(aic_model4)))
)

summary_stats$Best_AIC <- round(summary_stats$Best_AIC, 2)
summary_stats$Mean_AIC <- round(summary_stats$Mean_AIC, 2)
summary_stats$SD_AIC <- round(summary_stats$SD_AIC, 2)
summary_stats$Range_AIC <- round(summary_stats$Range_AIC, 2)

print(summary_stats)

# =============================================================================
# FINAL RECOMMENDATIONS
# =============================================================================

cat("\n=== FINAL RECOMMENDATIONS ===\n")

delta_threshold_strong <- 2
delta_threshold_weak <- 7

if(min(overall_comparison$Delta_AIC[-1]) > delta_threshold_strong) {
  cat("STRONG EVIDENCE: The best model has strong support (all other models ΔAICc > 2)\n")
} else if(min(overall_comparison$Delta_AIC[-1]) > delta_threshold_weak) {
  cat("MODERATE EVIDENCE: The best model has moderate support\n")
} else {
  cat("WEAK EVIDENCE: Multiple models have similar support - consider model averaging\n")
}

cat("\nRecommended model for further analysis:\n")
cat("Type:", best_overall$Model_Type, "\n")
cat("Predictor:", best_overall$Best_Predictor, "\n")
cat("AIC:", round(best_overall$Best_AIC, 2), "\n")
cat("Evidence ratio vs second best:", round(exp(0.5 * overall_comparison$Delta_AIC[2]), 1), ":1\n")

```
## Model comparison across ROIs
```{r}
# =============================================================================
# CORRECT AIC ANALYSIS - COMPARING MODELS WITHIN EACH ROI (NO MODEL 2)
# =============================================================================

# Load required library
library(dplyr)

# Function to compare models for a single ROI
compare_models_for_roi <- function(roi_name) {
  
  # Check if ROI exists in all model results
  if(!roi_name %in% names(model1$results) || 
     !roi_name %in% names(model3$results) || 
     !roi_name %in% names(model4$results)) {
    return(NULL)
  }
  
  # Extract models for this specific ROI
  model1_roi <- model1$results[[roi_name]]$model  # PC only
  model3_roi <- model3$results[[roi_name]]$model  # PC * OzDelta  
  model4_roi <- model4$results[[roi_name]]$model  # PC + OzDelta
  
  # Create comparison for this ROI (excluding Model 2)
  roi_comparison <- data.frame(
    ROI = roi_name,
    Model_Type = c("Model 1 (PC only)", 
                   "Model 3 (PC * OzDelta)", 
                   "Model 4 (PC + OzDelta)"),
    AIC = c(AIC(model1_roi), 
            AIC(model3_roi), 
            AIC(model4_roi)),
    stringsAsFactors = FALSE
  )
  
  # Add delta AIC and rankings within this ROI
  roi_comparison$Delta_AIC <- roi_comparison$AIC - min(roi_comparison$AIC)
  roi_comparison$Rank <- rank(roi_comparison$AIC)
  roi_comparison$AIC_Weight <- exp(-0.5 * roi_comparison$Delta_AIC) / 
                               sum(exp(-0.5 * roi_comparison$Delta_AIC))
  
  # Sort by AIC
  roi_comparison <- roi_comparison[order(roi_comparison$AIC), ]
  
  return(roi_comparison)
}

# =============================================================================
# ANALYSIS FOR EACH ROI
# =============================================================================

# Get ROI names that exist in all models
roi_names <- intersect(intersect(names(model1$results), names(model3$results)), 
                       names(model4$results))

cat("Analyzing", length(roi_names), "ROIs\n")

# Compare models for each ROI
all_roi_comparisons <- list()
for(roi in roi_names) {
  result <- compare_models_for_roi(roi)
  if(!is.null(result)) {
    all_roi_comparisons[[roi]] <- result
  }
}

# Combine all comparisons
combined_comparisons <- do.call(rbind, all_roi_comparisons)

cat("=== MODEL COMPARISON WITHIN EACH ROI ===\n")

# Summary: Which model wins most often?
best_models <- combined_comparisons[combined_comparisons$Rank == 1, ]
model_wins <- table(best_models$Model_Type)
cat("Model wins across all ROIs:\n")
print(model_wins)
cat("\n")

# Show best model for each ROI
cat("BEST MODEL FOR EACH ROI (first 10):\n")
best_by_roi <- combined_comparisons[combined_comparisons$Rank == 1, 
                                   c("ROI", "Model_Type", "AIC", "Delta_AIC")]
print(head(best_by_roi, 10))

# =============================================================================
# SUMMARY STATISTICS BY MODEL TYPE
# =============================================================================

cat("\n=== SUMMARY STATISTICS ACROSS ROIS ===\n")

# Calculate summary statistics manually
model_types <- unique(combined_comparisons$Model_Type)
summary_stats <- data.frame()

for(model_type in model_types) {
  model_data <- combined_comparisons[combined_comparisons$Model_Type == model_type, ]
  
  stats <- data.frame(
    Model_Type = model_type,
    N_ROIs = nrow(model_data),
    Times_Best = sum(model_data$Rank == 1),
    Mean_AIC = round(mean(model_data$AIC), 2),
    SD_AIC = round(sd(model_data$AIC), 2),
    Mean_Delta_AIC = round(mean(model_data$Delta_AIC), 2),
    Mean_Weight = round(mean(model_data$AIC_Weight), 3),
    stringsAsFactors = FALSE
  )
  
  summary_stats <- rbind(summary_stats, stats)
}

# Sort by times best, then by mean delta AIC
summary_stats <- summary_stats[order(-summary_stats$Times_Best, summary_stats$Mean_Delta_AIC), ]
print(summary_stats)

# =============================================================================
# TOP PERFORMING ROIS WITH DETAILED AIC DIFFERENCES
# =============================================================================

cat("\n=== TOP 5 BEST PERFORMING ROIS WITH AIC DIFFERENCES ===\n")

# Find best AIC for each ROI
best_aic_by_roi <- aggregate(AIC ~ ROI, data = combined_comparisons, FUN = min)
best_aic_by_roi <- best_aic_by_roi[order(best_aic_by_roi$AIC), ]
top5_rois <- head(best_aic_by_roi, 5)

for(i in 1:nrow(top5_rois)) {
  roi <- top5_rois$ROI[i]
  roi_details <- all_roi_comparisons[[roi]]
  
  cat(sprintf("\n%d. %s (Best AIC: %.2f)\n", i, roi, top5_rois$AIC[i]))
  
  # Extract AIC values for each model
  model1_aic <- roi_details$AIC[roi_details$Model_Type == "Model 1 (PC only)"]
  model3_aic <- roi_details$AIC[roi_details$Model_Type == "Model 3 (PC + OzDelta)"]
  model4_aic <- roi_details$AIC[roi_details$Model_Type == "Model 4 (PC * OzDelta)"]
  
  # Print model details with rankings
  print(roi_details[, c("Model_Type", "AIC", "Delta_AIC", "Rank")])
  
  # Calculate and print pairwise AIC differences
  cat("\nPairwise AIC Differences:\n")
  cat(sprintf("Model 3 vs Model 1: %.2f (", model1_aic - model3_aic))
  if(model1_aic - model3_aic > 2) {
    cat("Model 3 significantly better)\n")
  } else if(model3_aic - model1_aic > 2) {
    cat("Model 1 significantly better)\n")
  } else {
    cat("No strong preference)\n")
  }
  
  cat(sprintf("Model 4 vs Model 1: %.2f (", model1_aic - model4_aic))
  if(model1_aic - model4_aic > 2) {
    cat("Model 4 significantly better)\n")
  } else if(model4_aic - model1_aic > 2) {
    cat("Model 1 significantly better)\n")
  } else {
    cat("No strong preference)\n")
  }
  
  cat(sprintf("Model 4 vs Model 3: %.2f (", model3_aic - model4_aic))
  if(model3_aic - model4_aic > 2) {
    cat("Model 4 significantly better)\n")
  } else if(model4_aic - model3_aic > 2) {
    cat("Model 3 significantly better)\n")
  } else {
    cat("No strong preference)\n")
  }
  
  # Summary interpretation
  best_model <- roi_details$Model_Type[roi_details$Rank == 1]
  cat(sprintf("\nSummary: %s is best for %s\n", best_model, roi))
  
  # Evidence strength
  min_delta <- min(roi_details$Delta_AIC[roi_details$Rank > 1])
  if(min_delta > 10) {
    cat("Evidence: Very strong support for best model\n")
  } else if(min_delta > 4) {
    cat("Evidence: Strong support for best model\n")
  } else if(min_delta > 2) {
    cat("Evidence: Moderate support for best model\n")
  } else {
    cat("Evidence: Weak support - multiple models competitive\n")
  }
  
  cat(rep("-", 60), "\n")
}

# =============================================================================
# SUMMARY TABLE OF AIC DIFFERENCES FOR TOP ROIS
# =============================================================================

cat("\n=== SUMMARY TABLE: AIC DIFFERENCES FOR TOP 5 ROIS ===\n")

# Create summary table
top_roi_summary <- data.frame()

for(i in 1:nrow(top5_rois)) {
  roi <- top5_rois$ROI[i]
  roi_details <- all_roi_comparisons[[roi]]
  
  # Extract AIC values
  model1_aic <- roi_details$AIC[roi_details$Model_Type == "Model 1 (PC only)"]
  model3_aic <- roi_details$AIC[roi_details$Model_Type == "Model 3 (PC + OzDelta)"]
  model4_aic <- roi_details$AIC[roi_details$Model_Type == "Model 4 (PC * OzDelta)"]
  
  # Best model
  best_model <- roi_details$Model_Type[roi_details$Rank == 1]
  best_aic <- min(roi_details$AIC)
  
  # Create summary row
  summary_row <- data.frame(
    Rank = i,
    ROI = roi,
    Best_Model = best_model,
    Best_AIC = round(best_aic, 2),
    Model1_AIC = round(model1_aic, 2),
    Model3_AIC = round(model3_aic, 2),
    Model4_AIC = round(model4_aic, 2),
    Diff_M3_vs_M1 = round(model1_aic - model3_aic, 2),
    Diff_M4_vs_M1 = round(model1_aic - model4_aic, 2),
    Diff_M4_vs_M3 = round(model3_aic - model4_aic, 2),
    stringsAsFactors = FALSE
  )
  
  top_roi_summary <- rbind(top_roi_summary, summary_row)
}

print(top_roi_summary)

# =============================================================================
# INTERPRETATION OF DIFFERENCES
# =============================================================================

cat("\n=== INTERPRETATION GUIDE ===\n")
cat("AIC Difference Interpretation:\n")
cat("• Δ > 10: Very strong evidence against higher AIC model\n")
cat("• Δ > 4:  Strong evidence against higher AIC model\n")
cat("• Δ > 2:  Moderate evidence against higher AIC model\n")
cat("• Δ ≤ 2:  Weak evidence - models are competitive\n")

cat("\nColumn explanations:\n")
cat("• Diff_M3_vs_M1: Positive = Model 3 better, Negative = Model 1 better\n")
cat("• Diff_M4_vs_M1: Positive = Model 4 better, Negative = Model 1 better\n")
cat("• Diff_M4_vs_M3: Positive = Model 4 better, Negative = Model 3 better\n")

# Count significant improvements in top ROIs
cat("\n=== SIGNIFICANT IMPROVEMENTS IN TOP 5 ROIS ===\n")
cat("Model 3 significantly better than Model 1:", 
    sum(top_roi_summary$Diff_M3_vs_M1 > 2), "out of 5 ROIs\n")
cat("Model 4 significantly better than Model 1:", 
    sum(top_roi_summary$Diff_M4_vs_M1 > 2), "out of 5 ROIs\n")
cat("Model 4 significantly better than Model 3:", 
    sum(top_roi_summary$Diff_M4_vs_M3 > 2), "out of 5 ROIs\n")
```

## 
# Calculate the difference in slope of the relationship between PC and delirium when holding SWA at its 1st quartile vs. 3rd quartile value 
marginaleffects::slopes(model, variables = "PC", newdata = datagrid(SWA = iqr), hypothesis = "b2 - b1 = 0")
```{r}
response_var <- "bin_delirium"
covariates <- c("Diff_OzDeltaPow")
# Model 3: bin delirium ~ PC * Diff_OzDeltaPow
model3 <- run_glm_model(subject_data_ozdelta, pc_data, response_var, covariates) # nolint
# Model 4: bin delirium ~ PC + Diff_OzDeltaPow
model4 <- run_glm_model2(subject_data_ozdelta, pc_data, response_var, covariates) # nolint

# Version 1: Show slopes of PC variable at different OzDelta levels
marginaleffects::plot_slopes(
  model3$results$X17Networks_RH_DefaultB_PFCd_2$model,
  variables = "X17Networks_RH_DefaultB_PFCd_2",
  condition = "Diff_OzDeltaPow"
) +
  geom_hline(yintercept = 0, linetype = "dotted")

# Version 2: Show slopes of OzDelta at different PC levels
marginaleffects::plot_slopes(
  model3$results$X17Networks_RH_DefaultB_PFCd_2$model,
  variables = "Diff_OzDeltaPow",
  condition = "X17Networks_RH_DefaultB_PFCd_2"
) +
  geom_hline(yintercept = 0, linetype = "dotted")


# This will show only the difference test result
marginaleffects::slopes(
  model3$results$X17Networks_RH_DefaultB_PFCd_2$model,
  variables = "X17Networks_RH_DefaultB_PFCd_2",
  hypothesis = "b2 - b1 = 0",
  condition = list(
    Diff_OzDeltaPow = quantile(subject_data_ozdelta$Diff_OzDeltaPow, 
                               probs = c(0.25, 0.75), na.rm = TRUE)
  )
)
# alternative approach to determine the differences
marginaleffects::hypotheses(
  model3$results$X17Networks_RH_ContB_IPL_4$model,
  variables = "X17Networks_RH_ContB_IPL_4",
  hypothesis = "b2 - b1 = 0",
  condition = list(
    Diff_OzDeltaPow = quantile(subject_data_ozdelta$Diff_OzDeltaPow, 
                               probs = c(0.25, 0.75), na.rm = TRUE)
  )
)

# display the model effect of changing diff in OZ Delta on relationship between PC and delirium

## plotting the slope of the relationship between PC and delirium at different levels of SWA
m1 <-marginaleffects::plot_slopes(
  model3$results$X17Networks_RH_DefaultB_PFCd_2$model,
  variables = "X17Networks_RH_DefaultB_PFCd_2",
  condition = "Diff_OzDeltaPow"
) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "#6292c0") +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )
  print(m1)
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="Slope_PC_DefaultBPFCd2_DiffOzDelta"
  ggsave(paste0(filename, ".svg"),plot=m1, width = 10, height = 10)

```


Now that we've determined appropriate model fits with the AIC values
## Look at the marginal effects of the logistic regression model
rational is to convert the log ratio into a probability using
Marginalise and convert to probability scale
#results <- predict(model, newdata= dat, response = "response")
```{r}
# Extract the marginal effects
marginal_effects <- lapply(model3$results, function(res) {
  model <- res$model
  #marginalise and converts the log-odds into probabilities
  marginal_effects <- predict(model, type = "response")
  return(marginal_effects)
})
# this appears to be creating a probability distribution for each ROI?
View(marginal_effects)
marginal_effects_df <- as.data.frame(do.call(cbind, marginal_effects))
# Add column names (if not already present)
colnames(marginal_effects_df) <- names(marginal_effects)

prob <- marginal_effects$X17Networks_RH_DefaultB_PFCd_2
data_model_noNA <- na.omit(data_model_df) # 76 x 503 - this is the actual data the model was ran on

model <- model3$results$X17Networks_RH_DefaultB_PFCd_2$model
# Create an interaction plot
library(interactions)
# Get the model
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- model3$results[[model_name]]$model

# Create interaction plot with continuous moderator (no grouping)
i <- interact_plot(model,
  pred = "X17Networks_RH_DefaultB_PFCd_2",          # Predictor variable
  modx = "Diff_OzDeltaPow",                         # Moderator variable (continuous)
  #modx.values = NULL,                     # Use quartiles for moderator
  modx.values = "plus-minus",                       # Use actual values, not groups
  plot.points = TRUE,                               # Plot raw data points
  x.label = "PC", 
  y.label = "Predicted Probability",
  #colors = c("#1f77b4")                            # Single color since no grouping
) +
theme_minimal() +
theme(
  panel.grid.minor = element_blank(),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  legend.text = element_text(size = 12),
  legend.title = element_text(size = 14)
)

print(i)
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="Interaction Relationship"
  ggsave(paste0(filename, ".svg"),plot=i, width = 10, height = 10)


# Get the exact variable names
roi_name <- sig_interactions$label[1]
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- glm_model_ozdelta_pc$results[[model_name]]$model

# Find the PC variable name
pc_var <- names(model$model)[grep("^X", names(model$model))][1]

# Use the exact variable name
i <- interact_plot(model,
  pred = "X17Networks_RH_DefaultB_PFCd_2",                    # Use exact variable name
  modx = "Diff_OzDeltaPow",
  modx.values = "NULL", 
  plot.points = TRUE,
x.label = "PC",
y.label = "Predicted Probability",
colors = c("#7fb6cc","#1f77b4", "#f9a340")
) +
theme_minimal() +
theme(
  #panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),      # X-axis text size
  axis.text.y = element_text(size = 14),      # Y-axis text size
  axis.title.x = element_text(size = 16),     # X-axis title size
  axis.title.y = element_text(size = 16),     # Y-axis title size
  legend.text = element_text(size = 12),      # Legend text size
  legend.title = element_text(size = 14)  
)

print(i)

```

## Custom Interaction Plot - Fitted Model Curves Over Data
```{r}
plot_simple_logistic <- function(model_results, roi_name, data_df) {
  library(ggplot2)
  library(dplyr)
  
  # Get the specific model for the ROI
  model_name <- paste0("X", roi_name)
  model <- model_results$results[[model_name]]$model
  model_data <- model$model
  
  # Find the PC variable name
  pc_var <- names(model_data)[grep("^X", names(model_data))][1]
  
  # Create range for PC values
  pc_range <- seq(min(model_data[[pc_var]], na.rm = TRUE),
                  max(model_data[[pc_var]], na.rm = TRUE),
                  length.out = 100)
  
  # Create prediction data using all combinations of PC range and actual Oz Delta values
  pred_data_list <- list()
  for(i in 1:nrow(model_data)) {
    temp_data <- data.frame(
      PC = pc_range,
      Diff_OzDeltaPow = model_data$Diff_OzDeltaPow[i]
    )
    names(temp_data) <- c(pc_var, "Diff_OzDeltaPow")
    temp_data$predicted_prob <- predict(model, newdata = temp_data, type = "response")
    temp_data$subject_id <- i
    pred_data_list[[i]] <- temp_data
  }
  
  # Combine all prediction data
  pred_data <- do.call(rbind, pred_data_list)
  
  # Create the plot
  curve_plot <- ggplot() +
    # Add prediction curves for each subject (with transparency)
    geom_line(data = pred_data, 
             aes_string(x = pc_var, y = "predicted_prob", group = "subject_id"),
             alpha = 0.1, color = "#1f77b4") +
    # Add average prediction curve
    stat_summary(data = pred_data, 
                aes_string(x = pc_var, y = "predicted_prob"),
                fun = mean, geom = "line", color = "#1f77b4", size = 2) +
    # Add raw data points at bottom
    geom_point(data = model_data, 
              aes_string(x = pc_var, shape = "factor(response)"),
              y = 0.05, size = 2, alpha = 0.7, color = "black") +
    # Scales
    scale_shape_manual(values = c("0" = 1, "1" = 16),
                      name = "Delirium", labels = c("No", "Yes")) +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
    labs(
      x = "Participation Coefficient",
      y = "Predicted Probability",
      title = paste("Predicted Probability Curves:", roi_name),
      subtitle = "Individual curves (light) and average curve (dark) using actual Oz Delta values"
    ) +
    theme_minimal()
  
  return(curve_plot)
}

# Usage:
simple_plot <- plot_simple_logistic(glm_model_ozdelta_pc, sig_interactions$label[1], data_model_df)
print(simple_plot)

```
## simple logistic regression plot
```{r}
# Function to plot raw data with predicted probability curves overlaid
plot_raw_data_with_curves <- function(model_results, roi_name) {
  library(ggplot2)
  library(dplyr)
  
  # Get the specific model for the ROI
  model_name <- paste0("X", roi_name)
  model <- model_results$results[[model_name]]$model
  model_data <- model$model
  
  # Find the PC variable name
  pc_var <- names(model_data)[grep("^X", names(model_data))][1]
  
  # Create prediction grid for contour lines
  pc_range <- seq(min(model_data[[pc_var]], na.rm = TRUE),
                  max(model_data[[pc_var]], na.rm = TRUE),
                  length.out = 50)
  
  oz_range <- seq(min(model_data$Diff_OzDeltaPow, na.rm = TRUE),
                  max(model_data$Diff_OzDeltaPow, na.rm = TRUE),
                  length.out = 50)
  
  # Create grid for predictions
  pred_grid <- expand.grid(
    PC = pc_range,
    OzDelta = oz_range
  )
  names(pred_grid) <- c(pc_var, "Diff_OzDeltaPow")
  
  # Get predicted probabilities
  pred_grid$predicted_prob <- predict(model, newdata = pred_grid, type = "response")
  
  # Create the plot
  p <- ggplot() +
    # Add raw data points first (background)
    geom_point(data = model_data, 
              aes_string(x = pc_var, y = "Diff_OzDeltaPow", color = "factor(response)"),
              size = 3, alpha = 0.7) +
    # Add contour lines for predicted probability
    geom_contour(data = pred_grid, 
                aes_string(x = pc_var, y = "Diff_OzDeltaPow", z = "predicted_prob"),
                color = "black", size = 1, bins = 5) +
    # Add contour labels
    geom_contour(data = pred_grid, 
                aes_string(x = pc_var, y = "Diff_OzDeltaPow", z = "predicted_prob"),
                color = "white", size = 0.5, bins = 5) +
    # Color scale for data points
    scale_color_manual(values = c("0" = "#f9a340", "1" = "#7fb6cc"), 
                      name = "Delirium", labels = c("No", "Yes")) +
    # Labels
    labs(
      x = "Participation Coefficient",
      y = "Oz Delta Power Difference",
      title = paste("Raw Data with Predicted Probability Contours:", roi_name),
      subtitle = "Points = actual data, black lines = predicted probability contours"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.position = "right",
      panel.grid.major = element_blank(),  # Remove major gridlines
      panel.grid.minor = element_blank()   # Remove minor gridlines
    )
  return(p)
}

# Alternative version with specific probability contour lines
plot_raw_data_with_specific_curves <- function(model_results, roi_name, prob_levels = c(0.50,0.75)) {
  library(ggplot2)
  library(dplyr)
  
  # Get the specific model for the ROI
  model_name <- paste0("X", roi_name)
  model <- model_results$results[[model_name]]$model
  model_data <- model$model
  
  # Find the PC variable name
  pc_var <- names(model_data)[grep("^X", names(model_data))][1]
  
  # Create prediction grid
  pc_range <- seq(min(model_data[[pc_var]], na.rm = TRUE),
                  max(model_data[[pc_var]], na.rm = TRUE),
                  length.out = 100)
  
  oz_range <- seq(min(model_data$Diff_OzDeltaPow, na.rm = TRUE),
                  max(model_data$Diff_OzDeltaPow, na.rm = TRUE),
                  length.out = 100)
  
  pred_grid <- expand.grid(
    PC = pc_range,
    OzDelta = oz_range
  )
  names(pred_grid) <- c(pc_var, "Diff_OzDeltaPow")
  pred_grid$predicted_prob <- predict(model, newdata = pred_grid, type = "response")
  
  # Create the plot
  p <- ggplot() +
    # Raw data points
    geom_point(data = model_data, 
              aes_string(x = pc_var, y = "Diff_OzDeltaPow", color = "factor(response)"),
              size = 4, alpha = 0.8) +
    # Specific probability contour lines
    geom_contour(data = pred_grid, 
                aes_string(x = pc_var, y = "Diff_OzDeltaPow", z = "predicted_prob"),
                breaks = prob_levels, color = "black", size = 1.5) +
    # Color scale
    scale_color_manual(values =  c("0" = "#f9a340", "1" = "#7fb6cc"), 
                      name = "Delirium", labels = c("No", "Yes")) +
    labs(
      x = "Participation Coefficient",
      y = "Oz Delta Power Difference", 
      title = paste("PC vs Oz Delta with Probability Curves:", roi_name),
      subtitle = paste("Black lines show", paste(prob_levels*100, collapse = "%, "), "% predicted probability")
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.position = "right",
      panel.grid.major = element_blank(),  # Remove major gridlines
      panel.grid.minor = element_blank()   # Remove minor gridlines
    )
  return(p)
}

# Usage:
# Version 1: Multiple contour lines
raw_with_curves <- plot_raw_data_with_curves(glm_model_ozdelta_pc, sig_interactions$label[2])
print(raw_with_curves)

# Version 2: Specific probability levels (25%, 50%, 75%)
raw_with_specific <- plot_raw_data_with_specific_curves(glm_model_ozdelta_pc, sig_interactions$label[2])
print(raw_with_specific)
```








## Final Figure plots

# Custom interaction plot of fitted curve
```{r}
# Custom interaction plot using raw Diff_OzDeltaPow values
library(ggplot2)
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- glm_model_ozdelta_pc$results[[model_name]]$model
# Get model data
model_data <- model$model
pc_var <- "X17Networks_RH_DefaultB_PFCd_2"

# Create prediction grid with actual Oz Delta values
pc_range <- seq(min(model_data[[pc_var]], na.rm = TRUE),
                max(model_data[[pc_var]], na.rm = TRUE),
                length.out = 100)

# Use each subject's actual Oz Delta value for predictions
# Create prediction data using all combinations of PC range and actual Oz Delta values
pred_data_list <- list()
for(i in 1:nrow(model_data)) {
  temp_data <- data.frame(
    PC = pc_range,
    Diff_OzDeltaPow = model_data$Diff_OzDeltaPow[i]
  )
  names(temp_data)[1] <- pc_var
  temp_data$predicted_prob <- predict(model, newdata = temp_data, type = "response")
  pred_data_list[[i]] <- temp_data
}

pred_data <- do.call(rbind, pred_data_list)

# Create the plot (average prediction curve only)
custom_plot <- ggplot() +
  # Average prediction curve
  stat_summary(data = pred_data, 
               aes_string(x = pc_var, y = "predicted_prob"),
               fun = mean, geom = "line", color = "#6292c0", size = 2) +
  # Raw data points
geom_point(data = model_data, 
           aes_string(x = pc_var, y = "response", color = "factor(response)"),
           alpha = 0.6, size = 2) +
scale_color_manual(values = c("0" = "#ff7f0e", "1" = "#80b6cd"),
                   name = "Delirium", 
                   labels = c("No", "Yes")) +
  labs(
    x = "PC",
    y = "Predicted Probability",
    title = "Interaction Plot (Average Prediction Curve Only)"
  ) +
  theme_minimal() +
  theme(
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  )

print(custom_plot)
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="RawDiffOZDelta_Interaction2"
  ggsave(paste0(filename, ".svg"),plot=custom_plot, width = 10, height = 10)


```
## plotting of raw data
```{r}
# plot the raw data based model_data
raw_data_plot <- ggplot(model_data, aes_string(x = pc_var, y = "Diff_OzDeltaPow", color = "factor(response)")) +
  geom_point(size = 4, alpha = 0.8) +
  scale_color_manual(values = c("0" = "#f9a340", "1" = "#7fb6cc"), 
                     name = "Delirium", labels = c("No", "Yes")) +
  labs(
    x = "PC",
    y = "Oz Delta Power Difference",
    title = "Raw Log of Oz Delta Power vs PC Data"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    legend.position = "right",
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank()   # Remove minor gridlines
  )

print(raw_data_plot)

# Calculate predicted probabilities for the actual data points
model_data$predicted_prob <- predict(model, type = "response")

# Create the plot with predicted probabilities
prob_plot <- ggplot(model_data, aes_string(x = pc_var, y = "predicted_prob")) +
  geom_point(aes(color = factor(response)), alpha = 0.7, size = 3) +
  geom_smooth(method = "loess", se = TRUE, color = "#6292c0", size = 1.5) +
  scale_color_manual(values = c("0" = "#ff7f0e", "1" = "#80b6cd"),
                     name = "Delirium", 
                     labels = c("No", "Yes")) +
  labs(
    x = "PC",
    y = "Predicted Probability",
    title = "Predicted Probabilities from Logistic Regression Model"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 18, hjust = 0.5)
  ) +
  ylim(0, 1)

print(prob_plot)

setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="PredictedProbabilities_RawData_Interaction"
  ggsave(paste0(filename, ".svg"),plot=prob_plot, width = 10, height = 10)
## attempt to plot with the X axis being SWA instead of PC

# Create the plot with Diff_OzDeltaPow on x-axis
prob_plot_oz <- ggplot(model_data, aes(x = Diff_OzDeltaPow, y = predicted_prob)) +
  geom_point(aes(color = factor(response)), alpha = 0.7, size = 3) +
  geom_smooth(method = "loess", se = TRUE, color = "#6292c0", size = 1.5) +
  scale_color_manual(values = c("0" = "#ff7f0e", "1" = "#80b6cd"),
                     name = "Delirium", 
                     labels = c("No", "Yes")) +
  labs(
    x = "Diff_OzDeltaPow",
    y = "Predicted Probability",
    title = "Predicted Probabilities vs Oz Delta Power Difference"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 18, hjust = 0.5)
  ) +
  ylim(0, 1)

print(prob_plot_oz)
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="PredictedProbabilities_RawData_Interaction_OzDelta"
  ggsave(paste0(filename, ".svg"),plot=prob_plot_oz, width = 10, height = 10)

```


```{r}
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- glm_model_ozdelta_pc$results[[model_name]]$model
# Get model data
model_data <- model$model
pc_var <- "X17Networks_RH_DefaultB_PFCd_2"

# Create prediction grid with actual Oz Delta values
pc_range <- seq(min(model_data[[pc_var]], na.rm = TRUE),
                max(model_data[[pc_var]], na.rm = TRUE),
                length.out = 100)

# Calculate mean and standard deviation for Oz Delta
oz_mean <- mean(model_data$Diff_OzDeltaPow, na.rm = TRUE)
oz_sd <- sd(model_data$Diff_OzDeltaPow, na.rm = TRUE)

# Create prediction data for mean, +1SD, and -1SD
oz_levels <- c(oz_mean - oz_sd, oz_mean, oz_mean + oz_sd)
oz_labels <- c("Mean - 1SD", "Mean", "Mean + 1SD")

pred_data_combined <- data.frame()

for(j in 1:length(oz_levels)) {
  temp_data <- data.frame(
    PC = pc_range,
    Diff_OzDeltaPow = oz_levels[j]
  )
  names(temp_data)[1] <- pc_var
  temp_data$predicted_prob <- predict(model, newdata = temp_data, type = "response")
  temp_data$oz_level <- oz_labels[j]
  pred_data_combined <- rbind(pred_data_combined, temp_data)
}

# Create the plot with three lines
custom_plot <- ggplot() +
  # Prediction curves for mean ± 1SD
  geom_line(data = pred_data_combined, 
            aes_string(x = pc_var, y = "predicted_prob", color = "oz_level"),
            size = 2) +
  # Raw data points
  geom_point(data = model_data, 
             aes_string(x = pc_var, y = "response", shape = "factor(response)"),
             alpha = 0.6, size = 2, color = "black") +
  scale_color_manual(values = c("Mean - 1SD" = "#1f77b4", 
                               "Mean" = "#ff7f0e", 
                               "Mean + 1SD" = "#2ca02c"),
                     name = "Oz Delta Level") +
  scale_shape_manual(values = c("0" = 1, "1" = 16),
                     name = "Delirium", 
                     labels = c("No", "Yes")) +
  labs(
    x = "PC",
    y = "Predicted Probability",
    title = "Interaction Plot (Mean ± 1 Standard Deviation)"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  )

print(custom_plot)

setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="DiffOZDelta_Mean_SD_Interaction2"
  ggsave(paste0(filename, ".svg"),plot=custom_plot, width = 10, height = 10)


```






# log odds ratio plot for logistic regression model
```{r}
# Custom log-odds ratio plot using raw Diff_OzDeltaPow values
library(ggplot2)
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- glm_model_ozdelta_pc$results[[model_name]]$model
# Get model data
model_data <- model$model
pc_var <- "X17Networks_RH_DefaultB_PFCd_2"

# Create prediction grid with actual Oz Delta values
pc_range <- seq(min(model_data[[pc_var]], na.rm = TRUE),
                max(model_data[[pc_var]], na.rm = TRUE),
                length.out = 100)

# Use each subject's actual Oz Delta value for predictions
# Create prediction data using all combinations of PC range and actual Oz Delta values
pred_data_list <- list()
for(i in 1:nrow(model_data)) {
  temp_data <- data.frame(
  PC = pc_range,
  Diff_OzDeltaPow = model_data$Diff_OzDeltaPow[i]
  )
  names(temp_data)[1] <- pc_var
  pred_out <- predict(model, newdata = temp_data, type = "link", se.fit = TRUE)
  temp_data$predicted_prob <- pred_out$fit
  pred_data_list[[i]] <- temp_data
}

pred_data <- do.call(rbind, pred_data_list)

# Create the plot (average prediction curve only)
custom_plot <- ggplot() +
  # Average prediction curve
  stat_summary(data = pred_data, 
               aes_string(x = pc_var, y = "predicted_prob"),
               fun = mean, geom = "line", color = "#6292c0", size = 2) +
  # Raw data points
  labs(
    x = "PC",
    y = "Log-Odds",
    title = "Log-Odds Plot (Average Prediction Curve Only)"
  ) +
  theme_minimal() +
  theme(
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  )

print(custom_plot)
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="RawDiffOZDelta_LogOdds_Interaction"
  ggsave(paste0(filename, ".svg"),plot=custom_plot, width = 10, height = 10)


```
# plot IQ range
```{r}
# Custom interaction plot using raw Diff_OzDeltaPow values
library(ggplot2)
# Get the exact variable names
library(interactions)  # Ensure interact_plot is available

roi_name <- sig_interactions$label[1]
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- glm_model_ozdelta_pc$results[[model_name]]$model

# Find the PC variable name
pc_var <- names(model$model)[grep("^X", names(model$model))][1]

# Use the exact variable name
quartile_values <- quantile(
  model_data$Diff_OzDeltaPow,
  c(0.25, 0.50, 0.75),
  na.rm = TRUE
)
iqr_plot <- interact_plot(
  model,
  pred = "X17Networks_RH_DefaultB_PFCd_2",
  modx = "Diff_OzDeltaPow",
  modx.values = quartile_values,
  plot.points = TRUE,
  x.label = "PC",
  y.label = "Predicted Probability",
  colors = c("#7fb6cc", "#1f77b4", "#f9a340")
) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 14),      # X-axis text size
    axis.text.y = element_text(size = 14),      # Y-axis text size
    axis.title.x = element_text(size = 16),     # X-axis title size
    axis.title.y = element_text(size = 16),     # Y-axis title size
    legend.text = element_text(size = 12),      # Legend text size
    legend.title = element_text(size = 14)
  )

print(iqr_plot)

setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file
filename="DiffOZDelta_Interaction_IQ"
ggsave(paste0(filename, ".svg"),plot=iqr_plot, width = 10, height = 10)

```





# interaction plot + raw data plotted
```{r}
# Debug and fix the data plots with error checking

# Check if required columns exist

data_plot1 <- data_model_noNA[, c("response", "X17Networks_RH_DefaultB_PFCd_2", "Diff_OzDeltaPow")]
# Create binned version more carefully
data_plot1$OzDelta_binned <- cut(data_model_noNA$Diff_OzDeltaPow, 
                                breaks = quantile(data_model_noNA$Diff_OzDeltaPow, 
                                                 probs = c(0, 0.33, 0.67, 1), 
                                                 na.rm = TRUE), 
                                labels = c("Low", "Medium", "High"),
                                include.lowest = TRUE)

# Check binning worked
cat("Binning results:", table(data_plot1$OzDelta_binned), "\n")

# DATA-ONLY PLOT (Raw data scatter plot) - FIXED VERSION
data_plot <- ggplot(data_plot1, aes(x = X17Networks_RH_DefaultB_PFCd_2, 
                                   y = response)) + # nolint: indentation_linter.
  geom_point(aes(color = OzDelta_binned), 
             size = 3, alpha = 0.7, 
             position = position_jitter(height = 0.02, width = 0)) +  # Add jitter for visibility
  geom_smooth(aes(color = OzDelta_binned), 
              method = "glm", 
              method.args = list(family = "binomial"),
              se = TRUE, alpha = 0.2) +
  scale_color_manual(values = c("Low" = "#1f77b4", 
                               "Medium" = "#ff7f0e", 
                               "High" = "#2ca02c"),
                     na.value = "grey50") +  # Handle any NAs
  labs(
    x = "PC (X17Networks_RH_DefaultB_PFCd_2)",
    y = "Delirium Occurrence",
    color = "Diff_OzDeltaPow",
    title = "Raw Data: PC vs Delirium by Oz Delta Power"
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 16, hjust = 0.5)
  ) +
  scale_y_continuous(breaks = c(0, 1), labels = c("No Delirium", "Delirium"))

# Test plot creation
tryCatch({
  print(data_plot)
  cat("Binned plot created successfully\n")
}, error = function(e) {
  cat("Error creating binned plot:", e$message, "\n")
})

# CONTINUOUS VERSION - FIXED
data_plot_continuous <- ggplot(data_plot1, aes(x = X17Networks_RH_DefaultB_PFCd_2, 
                                              y = Diff_OzDeltaPow)) +
  geom_point(aes(color = Diff_OzDeltaPow), 
             size = 3, alpha = 0.7,
             position = position_jitter(height = 0.02, width = 0)) +  # Add jitter
  scale_color_gradient2(low = "#faa441", 
                        mid = "#6292c0", 
                        high = "#80b6cd",
                        midpoint = median(data_plot1$Diff_OzDeltaPow, na.rm = TRUE),
                        name = "Diff_OzDeltaPow") +
  labs(
    x = "PC (X17Networks_RH_DefaultB_PFCd_2)",
    y = "Diff in Oz Delta Power",
    title = "Raw Data: PC vs Continuous Oz Delta Power (Delirium)"
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    plot.title = element_text(size = 16, hjust = 0.5)
  ) 
  #scale_y_continuous(breaks = c(0, 1), labels = c("No Delirium", "Delirium"))

# Test plot creation
print(data_plot_continuous)
  ggsave("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/Data_Plot_PC_OzDelta_Continuous.svg", 
         plot = data_plot_continuous, 
         width = 12, height = 8, 
         dpi = 300, device = "svg")

# Check current working directory
cat("Current working directory:", getwd(), "\n")

# Set working directory with error checking
target_dir <- "/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/"

if(dir.exists(target_dir)) {
  setwd(target_dir)
  cat("Successfully changed to:", getwd(), "\n")
} else {
  cat("Directory does not exist:", target_dir, "\n")
  cat("Creating directory...\n")
  dir.create(target_dir, recursive = TRUE)
  setwd(target_dir)
}

# Save plots with error handling
tryCatch({
  ggsave("Data_Plot_PC_OzDelta_Binned.svg", 
         plot = data_plot, 
         width = 12, height = 8, 
         dpi = 300, device = "svg")
  cat("Binned plot saved successfully\n")
}, error = function(e) {
  cat("Error saving binned plot:", e$message, "\n")
})

tryCatch({

  cat("Continuous plot saved successfully\n")
}, error = function(e) {
  cat("Error saving continuous plot:", e$message, "\n")
})

# Also save as PNG as backup
tryCatch({
  ggsave("Data_Plot_PC_OzDelta_Binned.png", 
         plot = data_plot, 
         width = 12, height = 8, 
         dpi = 300)
  cat("Binned plot PNG saved successfully\n")
}, error = function(e) {
  cat("Error saving binned PNG:", e$message, "\n")
})

tryCatch({
  ggsave("Data_Plot_PC_OzDelta_Continuous.png", 
         plot = data_plot_continuous, 
         width = 12, height = 8, 
         dpi = 300)
  cat("Continuous plot PNG saved successfully\n")
}, error = function(e) {
  cat("Error saving continuous PNG:", e$message, "\n")
})

# List files to confirm saving
cat("Files in directory:\n")
cat(paste(list.files(pattern = "Data_Plot"), collapse = "\n"), "\n")

# try 3D static plot using scatterplot3d
library(scatterplot3d)

# Prepare data
plot_data <- data_model_noNA[complete.cases(data_model_noNA[c("X17Networks_RH_DefaultB_PFCd_2", "Diff_OzDeltaPow", "response")]), ]

# Get predicted probabilities for actual data
plot_data$predicted_prob <- predict(model, newdata = plot_data, type = "response")

# Create color vector based on actual delirium outcome
colors <- ifelse(plot_data$response == 1, "#f9a340", "#7fb6cc")

# Create 3D plot
#svg("3D_Static_Plot.svg", width = 12, height = 10)
s3d <- scatterplot3d(
  x = plot_data$X17Networks_RH_DefaultB_PFCd_2,
  y = plot_data$Diff_OzDeltaPow,
  z = plot_data$predicted_prob,
  color = colors,
  pch = 16,
  cex.symbols = 1.5,
  xlab = "Participation Coefficient",
  ylab = "Diff Oz Delta Power",
  zlab = "Predicted Probability",
  main = "3D Interaction: PC × Oz Delta Power → Delirium Risk",
  grid = TRUE,
  box = TRUE,
  angle = 45
)
print(s3d)
# Add legend
legend("topright", 
       legend = c("No Delirium", "Delirium"),
       col = c("#7fb6cc", "#f9a340"),
       pch = 16,
       cex = 1.2)
dev.off()


```





# for the avg pc_data
```{r}
# Extract the actual model object (adjust 'avg_pc' if needed)
model <- glm_model_ozdelta_avgpc$results$avg_pc$model
# Create an interaction plot
i <- interact_plot(
  model,
  pred = "avg_pc",
  modx = "Diff_OzDeltaPow",
  plot.points = TRUE,
  x.label = "PC",
  y.label = "Predicted Probability"
) +
  #scale_color_manual(values = c("#f9a340", "#fde725", "#7fb6cc")) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )
print(i)

```

Look at the top 10 predictors based on the proability distribution
```{r}
# Calculate the maximum value for each predictor
max_values <- apply(marginal_effects_df, 2, max, na.rm = TRUE)

# Sort the predictors by their maximum values
sorted_max_values <- sort(max_values, decreasing = TRUE)

# Get the top 10 predictors
top_10_predictors <- head(sorted_max_values, 10)

# Print the top 10 predictors
print(top_10_predictors)

# Convert to a data frame
top_10_df <- data.frame(
  Predictor = names(top_10_predictors),
  MaxValue = top_10_predictors
)
# Print each predictor on a new line
cat(paste(names(top_10_predictors), "=", round(top_10_predictors, 4), "\n"))

# Plot the top 10 predictors
ggplot(top_10_df, aes(x = reorder(Predictor, MaxValue), y = MaxValue)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 10 Predictors with Highest Probabilities",
    x = "Predictor",
    y = "Maximum Probability"
  ) +
  theme_minimal()
```
## Mean & SD for the diff Delta Power Post-op
```{r}

data_delta_power <- na.omit(subject_data_ozdelta)
#mean of the diff delta power
mean_postdiff_delta <- mean(oz_delta_orig$OzDeltaPow_Post)
print(mean_postdiff_delta)
#sd of the diff delta power
sd_postdiff_delta <- sd(oz_delta_orig$OzDeltaPow_Post)
print(sd_postdiff_delta)

mean_prediff_delta <- mean(oz_delta_orig$OzDeltaPow_Pre)
print(mean_prediff_delta)
sd_prediff_delta <- sd(oz_delta_orig$OzDeltaPow_Pre)
print(sd_prediff_delta)

#csv oz_delta into
write.csv(oz_delta_orig, "/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/Oz_deltapower_organised.csv")
```

## Plot the probability distribution for 1 regions
```{r}
# plot individual region
data_combine <- data.frame(
  Probability = marginal_effects$X17Networks_RH_DefaultB_PFCd_2,
  PC = data_model_noNA$X17Networks_RH_DefaultB_PFCd_2
  Diff_OzDeltaPow = data_model_noNA$Diff_OzDeltaPow
)

# Scatter plot with point color based on Probability
s1 <- ggplot(data_combine, aes(x = PC, y = Diff_OzDeltaPow, color = Probability)) +
  geom_point(size = 3) +  # Add points with size 3
  scale_color_gradient(low = "#0C7D79", high = "#C03B2B") +  # Define color gradient
  geom_smooth(method = "lm", se = TRUE, color = "black") +  # Add linear regression line
  labs(
    title = "Scatter Plot of PC X17Networks_RH_DefaultB_PFCd_2* Diff_OzDeltaPow",
    x = "PC for Significant ROI",
    y = "Diff OzDeltaPow",
    color = "Probability"
  ) +
  theme_minimal()+
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 14),
    axis.text = element_text(size =12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.position = "right",
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank()   # Remove minor gridlines
  )
    print(s1)
setwd("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Graph_Theory/schaef_400/lm_model/update_model/") 
  # Save the plot as an SVG file

filename="Prob_DiffOZdelta_interact_PC_ROI1"
  #ggsave(paste0(filename, ".svg"),plot=s1, width = 10, height = 10)

```





```{r}
# Get the specific model for a significant ROI
#roi_name <- 17Networks_RH_DefaultB_PFCd_2
model_name <- "X17Networks_RH_DefaultB_PFCd_2"
model <- glm_model_ozdelta_pc$results[[model_name]]$model

# Get the model data
model_data <- model$model

# Find the PC variable name
pc_var <- model_name

# Create a grid of values for predictions
pc_seq <- seq(min(model_data[[pc_var]], na.rm = TRUE), 
              max(model_data[[pc_var]], na.rm = TRUE), 
              length.out = 100)

# Choose specific values of oz_delta_diff to plot (quartiles)
oz_levels <- quantile(model_data$Diff_OzDeltaPow, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)

# Generate predicted probabilities
newdata <- expand.grid(pc_value = pc_seq,
                       oz_delta_diff = oz_levels)

# Set proper column names to match your model
names(newdata) <- c(pc_var, "Diff_OzDeltaPow")

newdata$predicted_prob <- predict(model, newdata = newdata, type = "response")
newdata$oz_label <- factor(round(newdata$Diff_OzDeltaPow, 3))  # for coloring

# Plot original data and overlay predicted logistic curves
library(ggplot2)
# Get the specific model for a significant ROI
roi_name <- sig_interactions$label[378]
model_name <- paste0("X", roi_name)
model <- glm_model_ozdelta_pc$results[[model_name]]$model

# Get the model data
model_data <- model$model

# Find the PC variable name
pc_var <- names(model_data)[grep("^X", names(model_data))][88]

# Create a grid of values for predictions
pc_seq <- seq(min(model_data[[pc_var]], na.rm = TRUE), 
              max(model_data[[pc_var]], na.rm = TRUE), 
              length.out = 100)

# Choose specific values of oz_delta_diff to plot (quartiles)
oz_levels <- quantile(model_data$Diff_OzDeltaPow, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)

# Generate predicted probabilities
newdata <- expand.grid(pc_value = pc_seq,
                       oz_delta_diff = oz_levels)

# Set proper column names to match your model
names(newdata) <- c(pc_var, "Diff_OzDeltaPow")

newdata$predicted_prob <- predict(model, newdata = newdata, type = "response")
newdata$oz_label <- factor(round(newdata$Diff_OzDeltaPow, 3))  # for coloring

# Plot original data and overlay predicted logistic curves
library(ggplot2)

ggplot() +
  # Original data points
  geom_jitter(data = model_data,
              aes_string(x = pc_var, y = "response", fill = "Diff_OzDeltaPow"),
              width = 0.01, height = 0.02, alpha = 0.6, size = 3, shape = 21) +
  # Predicted curves
  geom_line(data = newdata,
            aes_string(x = pc_var, y = "predicted_prob", color = "oz_label"),
            size = 1.5) +
  # Color scales
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#d73027",
                       name = "Oz Delta\n(Observed)") +
  scale_color_manual(values = c("#1f77b4", "#ff7f0e", "#2ca02c"),
                     name = "Oz Delta\n(Quartiles)") +
  labs(x = "Participation Coefficient",
       y = "Delirium (Observed & Predicted Probability)",
       title = paste("Logistic Curves for ROI:", roi_name)) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16)
)

```

```{r}
# Create proper overlay plot with model prediction contours
# First, create a finer grid for smooth contour lines
pc_fine <- seq(min(model_data[[pc_var]], na.rm = TRUE), 
               max(model_data[[pc_var]], na.rm = TRUE), 
               length.out = 50)
oz_fine <- seq(min(model_data$Diff_OzDeltaPow, na.rm = TRUE), 
               max(model_data$Diff_OzDeltaPow, na.rm = TRUE), 
               length.out = 50)

pred_grid <- expand.grid(pc_value = pc_fine, oz_value = oz_fine)
names(pred_grid) <- c(pc_var, "Diff_OzDeltaPow")

# Get predicted probabilities for the full grid
pred_grid$predicted_prob <- predict(model, newdata = pred_grid, type = "response")

# Create the overlay plot
overlay_plot <- ggplot() +
# Contour lines showing predicted probability
geom_contour(
  data = pred_grid,
  aes_string(x = pc_var, y = "Diff_OzDeltaPow", z = "predicted_prob"),
  color = "black", size = 0.8, alpha = 0.7
) +
# Original data points
geom_point(
  data = model_data,
  aes_string(x = pc_var, y = "Diff_OzDeltaPow", color = "factor(response)"),
  size = 3, alpha = 0.7
) +
# Color scale for observed delirium
scale_color_manual(
  values = c("0" = "#2166ac", "1" = "#d73027"),
  name = "Delirium\nObserved",
  labels = c("No", "Yes")
) +
labs(
  x = "Participation Coefficient",
  y = "Oz Delta Power",
  title = paste("Delirium Prediction Model with Contours:", roi_name),
  subtitle = "Contour lines show predicted probability of delirium"
) +
theme_minimal() +
theme(
  panel.grid = element_blank(),
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 14),
  legend.text = element_text(size = 12),
  legend.title = element_text(size = 12),
  plot.title = element_text(size = 16),
  plot.subtitle = element_text(size = 12)
)

print(overlay_plot)

# Set pc_var based on ROI name directly
pc_var <- paste0("PC_", roi_name)  # e.g., "PC_LeftHippocampus"

# Or if your column names have a different pattern:
pc_var <- paste0(roi_name, "_PC")  # e.g., "LeftHippocampus_PC"

# Alternative version with filled contours (heatmap style)
heatmap_plot <- ggplot() +
  # Filled contour (heatmap) showing predicted probability
  geom_contour_filled(data = pred_grid,
                      aes_string(x = pc_var, y = "Diff_OzDeltaPow", z = "predicted_prob"),
                      alpha = 0.6) +
  # Original data points on top
  geom_point(data = model_data,
             aes_string(x = pc_var, y = "Diff_OzDeltaPow", 
                       shape = "factor(response)", color = "factor(response)"),
             size = 4, stroke = 1.5) +
  # Custom colors and shapes
  scale_shape_manual(values = c("0" = 1, "1" = 19),  # hollow vs filled circles
                     name = "Delirium\nObserved",
                     labels = c("No", "Yes")) +
  scale_color_manual(values = c("0" = "white", "1" = "black"),
                     name = "Delirium\nObserved",
                     labels = c("No", "Yes")) +
  scale_fill_viridis_d(name = "Predicted\nProbability") +
  labs(x = "Participation Coefficient",
       y = "Oz Delta Power",
       title = paste("Delirium Prediction Heatmap:", roi_name)) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        plot.title = element_text(size = 16))

print(heatmap_plot)
```

```{r}
## Fitted Model Curve Plot - PC vs Delta Power
# Function to plot the fitted curve with PC on x-axis and delta power on y-axis
plot_fitted_curve_top75 <- function(model, model_data, pc_var_name, roi_label) {
  library(ggplot2)
  
  # Get the actual data used in the model
  plot_data <- model_data
  
  # Create a sequence of PC values for smooth curve
  pc_range <- seq(min(plot_data[[pc_var_name]], na.rm = TRUE), 
                  max(plot_data[[pc_var_name]], na.rm = TRUE), 
                  length.out = 100)
  
  # Use only the 75th percentile (top quartile) of delta power
  delta_75 <- quantile(plot_data$Diff_OzDeltaPow, 0.85, na.rm = TRUE)
  
  # Create prediction data for 75th percentile delta power
  pred_data <- data.frame(
    PC = pc_range,
    Diff_OzDeltaPow = delta_75
  )
  names(pred_data)[1] <- pc_var_name
  pred_data$predicted_prob <- predict(model, newdata = pred_data, type = "response")
  
  # Only keep points where predicted probability is >= 0.75
  pred_data_top <- pred_data[pred_data$predicted_prob >= 0.75, ]
  
  # Create the plot
  p <- ggplot(pred_data_top, aes_string(x = pc_var_name, y = "predicted_prob")) +
    geom_line(size = 1.5, color = "#2ca02c") +
    scale_y_continuous(limits = c(0.75, 1), labels = scales::percent_format()) +
    labs(
      x = "Participation Coefficient",
      y = "Predicted Probability of Delirium",
      title = paste("Fitted Model Curve (Top 75% Prob):", roi_label),
      subtitle = "Only showing predicted probability ≥ 75% at 75th percentile Delta Power"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
  
  return(p)
}
plot_fitted_curve_pc_delta <- function(model, model_data, pc_var_name, roi_label) {
  library(ggplot2)
  
  # Get the actual data used in the model
  plot_data <- model_data
  
  # Create a sequence of PC values for smooth curve
  pc_range <- seq(min(plot_data[[pc_var_name]], na.rm = TRUE), 
                  max(plot_data[[pc_var_name]], na.rm = TRUE), 
                  length.out = 100)
  
  # Create a sequence of delta power values for contour lines
  delta_range <- seq(min(plot_data$Diff_OzDeltaPow, na.rm = TRUE),
                     max(plot_data$Diff_OzDeltaPow, na.rm = TRUE),
                     length.out = 100)
  
  # Create prediction grid
  pred_grid <- expand.grid(
    PC = pc_range,
    DeltaPow = delta_range
  )
  names(pred_grid) <- c(pc_var_name, "Diff_OzDeltaPow")
  
  # Get predicted probabilities
  pred_grid$predicted_prob <- predict(model, newdata = pred_grid, type = "response")
  
  # Create the plot with contour lines showing probability
  p <- ggplot(pred_grid, aes_string(x = pc_var_name, y = "Diff_OzDeltaPow")) +
    geom_contour_filled(aes(z = predicted_prob), alpha = 0.7) +
    geom_contour(aes(z = predicted_prob), color = "white", size = 0.5) +
    scale_fill_viridis_d(name = "Predicted\nProbability", labels = scales::percent_format()) +
    labs(
      x = "Participation Coefficient",
      y = "Difference in Oz Delta Power",
      title = paste("Fitted Model Surface:", roi_label),
      subtitle = "Contours show predicted probability of delirium"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
  
  return(p)
}

# Alternative: Simple curve showing relationship at different delta power levels
plot_fitted_curve_simple <- function(model, model_data, pc_var_name, roi_label) {
  library(ggplot2)
  
  # Get the actual data used in the model
  plot_data <- model_data
  
  # Create a sequence of PC values for smooth curve
  pc_range <- seq(min(plot_data[[pc_var_name]], na.rm = TRUE), 
                  max(plot_data[[pc_var_name]], na.rm = TRUE), 
                  length.out = 100)
  
  # Use quantiles for different delta power levels
  delta_levels <- quantile(plot_data$Diff_OzDeltaPow, c(0.25, 0.5, 0.75), na.rm = TRUE)
  
  # Create prediction data for each delta level
  pred_list <- list()
  for(i in 1:length(delta_levels)) {
    pred_data <- data.frame(
      PC = pc_range,
      DeltaPow = delta_levels[i]
    )
    names(pred_data) <- c(pc_var_name, "Diff_OzDeltaPow")
    pred_data$predicted_prob <- predict(model, newdata = pred_data, type = "response")
    pred_data$delta_level <- names(delta_levels)[i]
    pred_list[[i]] <- pred_data
  }
  
  # Combine all predictions
  all_pred <- do.call(rbind, pred_list)
  
  # Create the plot
  p <- ggplot(all_pred, aes_string(x = pc_var_name, y = "predicted_prob", color = "delta_level")) +
    geom_line(size = 1.5) +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
    scale_color_manual(values = c("25%" = "#1f77b4", "50%" = "#ff7f0e", "75%" = "#2ca02c"),
                       name = "Delta Power\nLevel") +
    labs(
      x = "Participation Coefficient",
      y = "Predicted Probability of Delirium",
      title = paste("Fitted Model Curves:", roi_label),
      subtitle = "Lines show different levels of Oz Delta Power"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
  
  return(p)
}
```

# Plot fitted curve for the first significant ROI
```{r}
# Plot fitted curves for the top 10 predictors based on top_10_predictors
if (exists("top_10_predictors") && length(top_10_predictors) > 0) {
  for (roi_label in names(top_10_predictors)) {
    model_name <- paste0("X", roi_label)
    if (model_name %in% names(glm_model_ozdelta_pc$results)) {
      model <- glm_model_ozdelta_pc$results[[model_name]]$model
      model_data <- model$model
      pc_var_name <- names(model_data)[grep("^X", names(model_data))][1]
      if (!is.na(pc_var_name)) {
        simple_plot <- plot_fitted_curve_simple(model, model_data, pc_var_name, roi_label)
        print(simple_plot)
        ggsave(paste0("Fitted_Curve_TopPredictor_", roi_label, ".svg"),
               plot = simple_plot, width = 10, height = 6)
      }
    }
  }
} else {
  cat("No top predictors found to plot.\n")
}
```



## Running logistic regression with cognitive data
```{r}
# Assuming subject_data, data_fc, response_var, and covariates are defined
cog_data$tmtb
covariates <- c("tmtb")
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_cog_pc <- run_glm_model(cog_data, pc_data, response_var, covariates) # nolint
data_model_df = glm_model_cog_pc$data_model_df #that's the actual data that went into the model
View(glm_model_cogpc$coef_pval_df)

# Extract the interactions between ROI and covariates
interactions <- glm_model_cog_pc$coef_pval_df[grep(":", glm_model_cog_pc$coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)


# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)


# Get the significant odds ratio
interactions$sig_OddsRatio <- ifelse(interactions$PValue < 0.05, interactions$OddsRatio, 0)
data <- data.frame(label = interactions$label, SignificantCoefficient = interactions$sig_OddsRatio)

#plot_brain_data_oddsratio(data, "PC_ROI_cogtmtb_sigodds")

data2 <- data.frame(label = interactions$label, SignificantCoefficient = interactions$SignificantCoefficient)

plot_brain_data(data2, "PC_ROI_cogtmtb_sigcoeffs")
```


## Attempt to run glm function for il8_diff
```{r}
# Assuming subject_data, data_fc, response_var, and covariates are defined
covariates <- c("il8_diff")
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_il8pc <- run_glm_model(subject_data, pc_data, response_var, covariates) # nolint

# Extract the interactions between ROI and covariates
interactions <- glm_model_il8pc$coef_pval_df[grep(":", glm_model_il8pc $coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)


# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)

data2 <- data.frame(label = interactions$label, SignificantCoefficient = interactions$SignificantCoefficient)

plot_brain_data(data2, "PC_ROI_il8diff_sigcoeffs")

```

## Run Delirium ~ tmtb + diff_OzDeltaPow * PC
```{r}
run_glm_model_cog_ozdelta <- function(subject_data, pc_data, response_var, covariates) {
  library(dplyr)
  
  # Ensure the response variable and covariates exist in the subject_data
  if (!all(c(response_var, covariates) %in% colnames(subject_data))) {
    stop("Response variable or covariates not found in subject_data.")
  }
  
  # Initialize lists to store results
  results <- list()
  coef_pval_df <- data.frame()
  aic_list <- c()
  
  # Iterate through each ROI in pc_data
  for (roi in colnames(pc_data)) {
    # Add the current ROI to the subject_data
    subject_data$ROI <- pc_data[[roi]]
    
    # Define the formula for the logistic regression
    formula <- as.formula(paste(response_var, "~", paste(c(covariates, "ROI", "Diff_OzDeltaPow:ROI"), collapse = " + ")))
    
    # Fit the logistic regression model
    model <- glm(formula, data = subject_data, family = binomial(link = "logit"))
    
    # Extract coefficients and p-values
    coef_summary <- summary(model)$coefficients
    coef_df <- data.frame(
      ROI = roi,
      Variable = rownames(coef_summary),
      Beta = coef_summary[, "Estimate"],
      StdError = coef_summary[, "Std. Error"],
      ZValue = coef_summary[, "z value"],
      PValue = coef_summary[, "Pr(>|z|)"]
    )
    
    # Append to the results
    coef_pval_df <- rbind(coef_pval_df, coef_df)
    aic_list <- c(aic_list, AIC(model))
    
    # Store the model in the results list
    results[[roi]] <- model
  }
  
  # Return the results as a list
  return(list(results = results, coef_pval_df = coef_pval_df, aic_list = aic_list))
}
```

```{r}
subject_data_ozdelta_cog <- merge(subject_data_ozdelta, cog_data[, c("patientid", "tmta", "tmtb", "tmt_diff")], by = "patientid", all.x = TRUE)
# Run the logistic regression model
 Define covariates and response variable
covariates <- c("tmtb", "Diff_OzDeltaPow")
response_var <- "bin_delirium"

# Run the logistic regression model
glm_model_cog_ozdelta_pc <- run_glm_model_cog_ozdelta(subject_data_ozdelta_cog, pc_data, response_var, covariates)

# View the results
#View(glm_model_cog_ozdelta_pc$coef_pval_df)

# Extract the interactions between ROI and covariates
interactions <- glm_model_cog_ozdelta_pc$coef_pval_df[grep(":", glm_model_cog_ozdelta_pc$coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)

colnames(interactions)[colnames(interactions) == "ROI"] <- "label"


# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)

data2 <- data.frame(label = interactions$label, SignificantCoefficient = interactions$SignificantCoefficient)

plot_brain_data(data2, "PC_ROI_il8diff_cogtmtb_sigcoeffs")




```






Note: Running the Log Reg. with FC (ignore for now)
## Relationship between ACh and LC FC and Delirium outcome 
```{r}
# load in FC data
fc_lc_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/avg_LC_fc_400_all_subs.csv", header = FALSE)
fc_nbm_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/avg_nbm_fc_400_all_subs.csv", header = FALSE)
fc_ppn_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Functional_Connectivity/AAS_nuclei/avg_ppn_fc_400_all_subs.csv", header = FALSE)

## Clinical Data
subject_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/subject_data_IL810_mri.csv")
# cog data 
cog_data <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/mri_scanner_cog.csv")
#load in the ROI names
roi_names <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/code/schaef_400roi_names.csv", header = FALSE)
## oz delta data
oz_delta_orig <- read.csv("/Users/ntaylor/Library/CloudStorage/OneDrive-TheUniversityofSydney(Staff)/Postdoc_Rob/Analysis/Demographic_Data/Oz_deltapower.csv")
# Remove "_" in subject names
oz_delta_orig$Sub <- gsub("_", "", oz_delta_orig$Sub)
# Filter for MRI subjects
oz_delta <- oz_delta_orig[oz_delta_orig$Sub %in% subject_data$patientid, ]
# Merge oz_delta with subject_data based on patientid
subject_data_ozdelta <- merge(subject_data, oz_delta, by.x = "patientid", by.y = "Sub", all.x = TRUE)

```




## Running the glm on the LC FC data - simple model, no covariates

## Take the exponentiate of the beta coefficients to convert to an odds ratio
Odds Ratio : The odds ratio is the exponentiated coefficient. It is the ratio of the odds of the event occurring in the exposed group to the odds of the event occurring in the unexposed group. An odds ratio of 1 indicates that the event is equally likely in both groups. An odds ratio greater than 1 indicates that the event is more likely in the exposed group, while an odds ratio less than 1 indicates that the event is less likely in the exposed group.
Interpretaiton - if B >0, then e^B >1, means that the functional connections increases, the odds of delirium increase (if in the inverse, increase in FC is associated with decrease odds of delirium)

```{r}
fc_data <- fc_lc_data
#rename colum names in pc_data with roi_names
colnames(fc_data) <- roi_names$V1[1:400]
# Assuming subject_data, data_fc, response_var, and no covariate
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_lc_fc <- run_glm_simple_model(subject_data, fc_data, response_var) # nolint

View(glm_modelcoeffs <- glm_model_lc_fc$coef_pval_df)
```

# get the above significant coeffs and plot them on the brain
```{r}
glm_model_interest <- glm_model_lc_fc #change to model of interest
#coefs and pvals from model
coeffs <- glm_model_interest$coef_pval_df[grep("^X", glm_model_interest$coef_pval_df$Variable), ]
# Add a column to interactions to indicate significant coefficients
coeffs$SignificantCoefficient <- ifelse(coeffs$PValue < 0.05, coeffs$Beta, 0)
#add a column that is Variable 
coeffs$label <- sub(":.*", "", coeffs$Variable)
coeffs$label <- sub("X", "", coeffs$label)

# Extract significant interactions
sig_roi_simple_glm <- coeffs[coeffs$PValue < 0.05, ]

coeffs$sig_OddsRatio <- ifelse(coeffs$PValue < 0.05, coeffs$OddsRatio, 0)

View(sig_roi_simple_glm)
View(coeffs)

# prepare data for plotting on brain using
data_lc_fc_sigodds <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$sig_OddsRatio)
plot_brain_data_oddsratio(data_lc_fc_sigodds, "lc_fc_sigodds_simpleglm")
#modify - because the 44 sig_OddsRatio is hard to visualise others
data_lc_fc_sigodds <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$sig_OddsRatio)



data_lc_fc_sigcoefs <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$SignificantCoefficient)
plot_brain_data(data_lc_fc_sigcoefs, "lc_fc_sigbeta_simpleglm")

```

```{r}
#remove the sig_OddsRatio > 40 & replace it with 0
data_lc_fc_sigodds$SignificantCoefficient[data_lc_fc_sigodds$SignificantCoefficient > 40] <- 0

plot_brain_data_oddsratio(data_lc_fc_sigodds, "lc_fc_sigodds_simpleglm_remove40")
```
## need to check the AIC for each ROI in the model
```{r}
View(glm_model_lc_fc$aic_list)

#plot the AIC values
aic_values <- unlist(glm_model_lc_fc$aic_list)
barplot(aic_values, main="AIC values for each ROI", xlab="ROI", ylab="AIC")
# Convert AIC values to a data frame for plotting
aic_df <- data.frame(ROI = names(aic_values), AIC = aic_values)

# Find the best performing region (lowest AIC)
best_region <- aic_df[which.min(aic_df$AIC), ]

# Plot the AIC values
ggplot(aic_df, aes(x = reorder(ROI, AIC), y = AIC)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(AIC, 2)), hjust = -0.3, size = 3) +
  geom_point(data = best_region, aes(x = ROI, y = AIC), color = "red", size = 3) +
  coord_flip() +
  labs(title = "AIC Values for Each ROI", x = "ROI", y = "AIC") +
  theme_minimal()

# Print the best performing region
print(best_region)

# is there a significant difference in the AIC values across regions?


```



```{r}
# Fit the logistic regression model
glm_model <- glm(subject_data$bin_delirium ~ ., data = fc_data, family = binomial)

# Calculate VIF for all predictors
vif_values <- vif(glm_model)

# Print the VIF values
print(vif_values)

# Identify predictors with high VIF
high_vif <- names(vif_values[vif_values > 5])
print(high_vif)
```


## Running Simple Log. Reg on FC ACh 
```{r}
fc_data <- fc_nbm_data
#rename colum names in pc_data with roi_names
colnames(fc_data) <- roi_names$V1[1:400]
# Assuming subject_data, data_fc, response_var, and no covariate
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_nbm_fc <- run_glm_simple_model(subject_data, fc_data, response_var) # nolint

View(glm_modelcoeffs <- glm_model_nbm_fc$coef_pval_df)
```

# get the above significant coeffs and plot them on the brain
```{r}
glm_model_interest <- glm_model_nbm_fc #change to model of interest
#coefs and pvals from model
coeffs <- glm_model_interest$coef_pval_df[grep("^X", glm_model_interest$coef_pval_df$Variable), ]
# Add a column to interactions to indicate significant coefficients
coeffs$SignificantCoefficient <- ifelse(coeffs$PValue < 0.05, coeffs$Beta, 0)
#add a column that is Variable 
coeffs$label <- sub(":.*", "", coeffs$Variable)
coeffs$label <- sub("X", "", coeffs$label)

# Extract significant interactions
sig_roi_simple_glm <- coeffs[coeffs$PValue < 0.05, ]

coeffs$sig_OddsRatio <- ifelse(coeffs$PValue < 0.05, coeffs$OddsRatio, 0)

View(sig_roi_simple_glm)
View(coeffs)

# prepare data for plotting on brain using
data_nbm_fc_sigodds <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$sig_OddsRatio)
plot_brain_data_oddsratio(data_nbm_fc_sigodds, "nbm_fc_sigodds_simpleglm")
# sig beta coeffs plotted
data_nbm_fc_sigcoefs <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$SignificantCoefficient)
plot_brain_data(data_nbm_fc_sigcoefs, "nbm_fc_sigbeta_simpleglm")

```

## need to check the AIC for each ROI in the model
```{r}
View(glm_model_lc_fc$aic_list)

#plot the AIC values
aic_values <- unlist(glm_model_lc_fc$aic_list)
barplot(aic_values, main="AIC values for each ROI", xlab="ROI", ylab="AIC")
# Convert AIC values to a data frame for plotting
aic_df <- data.frame(ROI = names(aic_values), AIC = aic_values)

# Find the best performing region (lowest AIC)
best_region <- aic_df[which.min(aic_df$AIC), ]

# Plot the AIC values
ggplot(aic_df, aes(x = reorder(ROI, AIC), y = AIC)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(AIC, 2)), hjust = -0.3, size = 3) +
  geom_point(data = best_region, aes(x = ROI, y = AIC), color = "red", size = 3) +
  coord_flip() +
  labs(title = "AIC Values for Each ROI", x = "ROI", y = "AIC") +
  theme_minimal()

# Print the best performing region
print(best_region)

# is there a significant difference in the AIC values across regions?


```

## Running Simple Log. Reg on FC PPN
```{r}
fc_data <- fc_ppn_data
#rename colum names in pc_data with roi_names
colnames(fc_data) <- roi_names$V1[1:400]
# Assuming subject_data, data_fc, response_var, and no covariate
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_ppn_fc <- run_glm_simple_model(subject_data, fc_data, response_var) # nolint

View(glm_model_ppn_fc$coef_pval_df)
```

# get the above significant coeffs and plot them on the brain
```{r}
glm_model_interest <- glm_model_ppn_fc #change to model of interest
#coefs and pvals from model
coeffs <- glm_model_interest$coef_pval_df[grep("^X", glm_model_interest$coef_pval_df$Variable), ]
# Add a column to interactions to indicate significant coefficients
coeffs$SignificantCoefficient <- ifelse(coeffs$PValue < 0.05, coeffs$Beta, 0)
#add a column that is Variable 
coeffs$label <- sub(":.*", "", coeffs$Variable)
coeffs$label <- sub("X", "", coeffs$label)

# Extract significant interactions
sig_roi_simple_glm <- coeffs[coeffs$PValue < 0.05, ]

coeffs$sig_OddsRatio <- ifelse(coeffs$PValue < 0.05, coeffs$OddsRatio, 0)

View(sig_roi_simple_glm)
View(coeffs)

# prepare data for plotting on brain using
data_ppn_fc_sigodds <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$sig_OddsRatio)
plot_brain_data(data_ppn_fc_sigodds, "ppn_fc_sigodds_simpleglm")
# sig beta coeffs plotted
data_ppn_fc_sigcoefs <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$SignificantCoefficient)
plot_brain_data(data_ppn_fc_sigcoefs, "ppn_fc_sigbeta_simpleglm")

```
# attempt to look at the marginal effects of the model
```{r}
# Assuming subject_data, data_fc, response_var, and covariates are defined
glm_model_interest <- glm_model_ppn_fc #change to model of interest


region_of_interest <- "X17Networks_LH_VisCent_ExStr_1"
exists <- region_of_interest %in% names(glm_model_interest$results$X17Networks_LH_VisCent_ExStr_1$marginal_effects)

# Print whether it exists
print(exists)

# Extract the marginal effects for each variable
marginal_effects <- lapply(glm_model_interest$results, function(res) {
  model <- res$model
  # Calculate marginal effects (predicted probabilities)
  marginal_effects <- predict(model, type = "response")
  return(marginal_effects)
})

# Combine marginal effects into a data frame for easier visualization
marginal_effects_df <- do.call(cbind, marginal_effects)
colnames(marginal_effects_df) <- names(glm_model_interest$results)

# View the marginal effects
View(marginal_effects_df)


```

## Running GLM on FC ACh and relevant cognitive scores from TMT-B

```{r}
fc_data <- fc_nbm_data
#rename colum names in pc_data with roi_names
colnames(fc_data) <- roi_names$V1[1:400]
# Assuming subject_data, data_fc, response_var, and covariates are defined
#covariates <- c("il8_diff")
covariates <- c("tmt_diff")
response_var <- "bin_delirium"

# Run the logistic regression model
glm_model_nbm_fc_cog <- run_glm_model(cog_data, fc_data, response_var, covariates) # nolint
View(glm_model_nbm_fc_cog$coef_pval_df)

glm_model_interest <- glm_model_nbm_fc_cog #change to model of interest

#coefs and pvals from model - interactions only
coeffs <- glm_model_interest$coef_pval_df[grep(":", glm_model_interest$coef_pval_df$Variable), ]
# Add a column to interactions to indicate significant coefficients
coeffs$SignificantCoefficient <- ifelse(coeffs$PValue < 0.05, coeffs$Beta, 0)
#add a column that is Variable 
coeffs$label <- sub(":.*", "", coeffs$Variable)
coeffs$label <- sub("X", "", coeffs$label)

# Extract significant interactions
sig_roi_simple_glm <- coeffs[coeffs$PValue < 0.05, ]

coeffs$sig_OddsRatio <- ifelse(coeffs$PValue < 0.05, coeffs$OddsRatio, 0)

View(sig_roi_simple_glm)
View(coeffs)
```
# given that there are PValues = 0, need to check for multicollinearity
```{r}
library(car)

# Example: Loop through each variable in the model
vif_results <- lapply(glm_model_nbm_fc_cog$results, function(res) {
  model <- res$model
  vif_values <- vif(model)
  return(vif_values)
})

# Combine VIF results into a data frame
vif_df <- do.call(rbind, lapply(names(vif_results), function(var) {
  vif_values <- vif_results[[var]]
  if (!is.na(vif_values)) {
    data.frame(Variable = names(vif_values), VIF = vif_values, Predictor = var)
  } else {
    NULL
  }
}))

# View the VIF results
View(vif_df)
```

```{r}

# prepare data for plotting on brain using
data_nbm_fc_sigodds <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$sig_OddsRatio)
plot_brain_data(data_nbm_fc_sigodds, "ppn_fc_sigodds_simpleglm")
# sig beta coeffs plotted
data_nbm_fc_sigcoefs <- data.frame(label = coeffs$label, SignificantCoefficient = coeffs$SignificantCoefficient)
plot_brain_data(data_lc_fc_sigcoefs, "ppn_fc_sigbeta_simpleglm")

```

## Variance Inflation Factor (VIF)
What It Does: VIF measures how much the variance of an estimated regression coefficient increases if your predictors are correlated.
How to Use It: Calculate VIF for each predictor. A common rule of thumb is that a VIF value above 5 or 10 suggests problematic multicollinearity.

## Avg. FC LC, nbM, PPN in the model

```{r}
# Fit the logistic regression model for bin_delirium ~ {avg. FC LC} + {avg. FC nbM} + {avg. FC PPN} 
avg_fc_lc <- rowMeans(fc_lc_data, na.rm = TRUE)
avg_fc_nbm <- rowMeans(fc_nbm_data, na.rm = TRUE)
avg_fc_ppn <- rowMeans(fc_ppn_data, na.rm = TRUE)

avg_fc_glm_model <- glm(subject_data$bin_delirium ~  avg_fc_lc + avg_fc_nbm + avg_fc_ppn, family = binomial)
summary_avg_fc_glm_model <- summary(avg_fc_glm_model)

# get the beta coeffs and pvals
coef_table <- summary_avg_fc_glm_model$coefficients
coef_pval_df <- data.frame(
  Variable = rownames(coef_table),
  Beta = coef_table[, "Estimate"],
  PValue = coef_table[, "Pr(>|z|)"],
  OddsRatio = exp(coef_table[, "Estimate"]) # Calculate odds ratio
)
View(coef_pval_df)

```


## Relationship between FC for LC, nbM and PPN * IL-8
```{r}
# Assuming subject_data, data_fc, response_var, and covariates are defined
fc_data <- fc_lc_data
colnames(fc_data) <- roi_names$V1[1:400]
covariates <- c("il8_diff")
response_var <- "bin_delirium"
# Run the logistic regression model
glm_model_lcfc_il8diff <- run_glm_model(subject_data, fc_data, response_var, covariates) # nolint
data_model_df = glm_model_lcfc_il8diff$data_model_df #that's the actual data that went into the model
View(glm_model_lcfc_il8diff$coef_pval_df)

# Extract the interactions between ROI and covariates
interactions <- glm_model_lcfc_il8diff$coef_pval_df[grep(":", glm_model_lcfc_il8diff$coef_pval_df$Variable), ]
View(interactions)

# Add a column to interactions to indicate significant coefficients
interactions$SignificantCoefficient <- ifelse(interactions$PValue < 0.05, interactions$Beta, 0)
#add a column that is Variable 
interactions$label <- sub(":.*", "", interactions$Variable)
interactions$label <- sub("X", "", interactions$label)


# Extract significant interactions
sig_interactions <- interactions[interactions$PValue < 0.05, ]
View(sig_interactions)


# Get the significant odds ratio
interactions$sig_OddsRatio <- ifelse(interactions$PValue < 0.05, interactions$OddsRatio, 0)
data <- data.frame(label = interactions$label, SignificantCoefficient = interactions$sig_OddsRatio)

plot_brain_data_oddsratio(data, "LCFC_il8diff_sigodds")

data2 <- data.frame(label = interactions$label, SignificantCoefficient = interactions$SignificantCoefficient)
plot_brain_data(data2, "LCFC_il8diff_sigcoefs")
```